```{r setup, include=F, echo = F}
library(tidyverse)
library(broom)
library(ggplot2)
library(lubridate)
library(RColorBrewer)
library(here)
library(reticulate)
library(patchwork)
library(snotelr)
library(zoo)
library(dataRetrieval)
library(cpop)

py_require("pip")
py_require("hydroserverpy")
py_require("datetime")
use_python("/Library/Frameworks/Python.framework/Versions/3.14/bin/python3", required = T)

knitr::opts_knit$set(root.dir = here("outputs", "htmls"))
```

## Overview

This script performs initial data cleanining, visualization, and basic analysis for five test sites in the Logan River Observatory network. These sites include: 

- Logan River at Franklin Basin (discharge, water temp)
- Logan River at Main Street (discharge, water temp)
- Logan River at Mendon (discharge, water temp)
- Franklin Basin Climate Station (Precipitation, Air temp)
- Logan River Golf Course Climate Station (Precipitation, Air temp)

Data is called directly from the LRO Hydroserver API (https://lro.hydroserver.org/browse)

This script performs the following actions: 

1. Read in and clean data files
  + Call Metadata for all sites in the LRO network
  + Filter out desired sites
  + Import data using a for loop
2. Plotting and basic analysis
  + Discharge
  + Stream Temperature

## I. Import Data

### A. LRO Data

First, Import necessary packages; establish connection to Hydroserver

```{python}
import hydroserverpy
import datetime
import pandas as pd

from datetime import datetime

from hydroserverpy import HydroServer

#initialize Hydroserver Connection with login Credentials **NOTE** you may need to update these to your Hydroserver login info

hs_api = HydroServer(host = 'https://lro.hydroserver.org', email = 'ggianniny@gmail.com',password='syxqyz-Zeqpor-7naxki')

#Load info for all datastreams:

datastreams = hs_api.datastreams.list(fetch_all = True)

```

#### i. Call metadata for all "datastreams" (i.e. all parameters at all sites) in the LRO network

Setup empty data frame to recieve data:

```{python}
df = pd.DataFrame(data = None, index = None, columns = ['uid', 'name', 'property','propCode','unit', 'medium', 'procLevel']) #initialize empty dataframe (outside loop)
```

Read in data with a for-loop (NOTE - this will take a few minutes to run)

```{python}
for i in range(0, len(datastreams.items)):
        uid = str(datastreams.items[i].uid) #extract UID
        name = str(datastreams.items[i].thing.name) #extract site name
        prop = str(datastreams.items[i].observed_property.name) #extract measured property
        propCode = str(datastreams.items[i].observed_property.code) #extract measured property code
        unit = str(datastreams.items[i].unit.name) #extract unit
        medium = str(datastreams.items[i].sampled_medium) #extract sampled medium
        procLevel = str(datastreams.items[i].processing_level.definition) #extract processing level
        siteInfo = pd.DataFrame(data = {'uid':[uid], 'name':[name], 'property':[prop], 'propCode':[propCode], 'unit':[unit],'medium':[medium], 'procLevel':[procLevel]}) #convert uid, name, and unit into dataframe 
        df = pd.concat([df, siteInfo]) #add siteInfo to df
        print("working on"+" "+str(i)) #print current site for progress tracking
```

#### ii. Filter out desired sites/parameters

First - convert to r, filter out quality controlled data, create reference DF's

```{r}
r_df <- py_to_r(py$df)%>% #converting to R for easier wrangling
  #filter(procLevel == "Quality controlled data"|procLevel == "Derived products")%>% #extract quality controlled and derived product data only
  arrange(name, property)%>% #sort by name and unit for easier viewing
  mutate(name = as.factor(name), 
         property = if_else(property == "Temperature", paste(medium, property, sep = " "), property), 
         property = if_else(str_detect(propCode, "WaterYr"), paste(property, "WaterYr", sep = " "), property),
         property = if_else(str_detect(propCode, "Discharge"), propCode, property),
         property = if_else(unit=="cubic feet per second", paste(property, "_cfs", sep =""), property),
         property = as.factor(property))%>%
  mutate(siteNum = as.numeric(name), 
         propNum = as.numeric(property))

nameRef <- r_df%>%
  select(name, siteNum)%>%
  distinct()%>% #Create dataframe with name references
  arrange(name)

propRef <- r_df%>%
  select(property, propNum)%>%
  distinct()%>%
  arrange(property)#Creste dataframe with property references
```

Next: view reference dataframes, write down site numbers and property numbers for export 

View name references:

```{r}
View(nameRef)
```

Enter desired site numbers below: 

```{r}
soi <- c(7,8,15,18, 19,20,21, 22, 23,24,26, 27, 34) #ENTER DESIRED SITE NUMBERS HERE inside the parentheses. e.g. if you want sites 1, 3, and 19, the code should read "soi <- c(1, 3, 19)"

#check correct sites selected:
nameRef[soi,]
```

View property references:

```{r}
View(propRef)
```

Enter desired property numbers below:

```{r}
poi <- c(1,6,12,37) #ENTER DESIRED PROPERTY NUMBERS HERE inside the parentheses. e.g. if you want properties 8, 17, and 21, the code should read "soi <- c(8, 17, 21)"

#check correct properties selected:
propRef[poi,]
```

Extract desired metadata:

```{r}
#extract desired datastreams:
toDownload <- r_df %>%
  filter(
    siteNum %in% soi&propNum %in% poi
  )%>% #extract sites/properties listed above
  mutate(siteProp = paste(name, property, sep =" "))%>% #add site + property column
  mutate(procLevelCode = case_when(
    procLevel == "Quality controlled data"~1, 
    procLevel == "Provisional data" ~2, 
    procLevel == "Raw data" ~3
  ))%>%
  arrange(name, property, procLevelCode)%>%
  mutate(dups = duplicated(siteProp))%>% #flag duplicated site+properties
  filter(!(dups == T & (procLevel == "Raw data"|procLevel == "Provisional data"))) #remove raw data datastreams for all sites that have quality controlled or provisional data

#Check to ensure all sites/data are present: 
#View(toDownload)

#convert back to Python for download: 

pyDownload <- r_to_py(toDownload)
```


#### iii. Import data using a for loop

Setup:

```{python}
#Read "to download" dataframe into Python environment
pyDownload = pd.DataFrame(r.pyDownload)

#initialize empty dataframe to receive data: 

fullData = pd.DataFrame(data = None, index = None, columns = ['phenomenon_time', 'result', 'site','param', 'unit']) #initialize empty dataframe (outside loop)
```


Read in all data using a For loop - NOTE, this will take a while to run (~2-4 min per site)

```{python}
for var in range(0, len(pyDownload["uid"])): #iterate on "i" over # rows in the pyDownload DF
  print("started"+" "+pyDownload.iloc[var,1]) #print current site name for progress tracking
  data = hs_api.datastreams.get_observations(fetch_all = True, uid = pyDownload.iloc[var,0]).dataframe #read in data to temporary df
  data['site'] = pyDownload.iloc[var,1] #add site name
  data['param'] = pyDownload.iloc[var,2] #add parameter name
  data['unit'] = pyDownload.iloc[var,4] #add unit 
  fullData = pd.concat([fullData, data]) #add data to fullData df created in block above. 
  print("finished"+" "+pyDownload.iloc[var,1]) #print current site name for progress tracking
```

Convert full data file back to R for additional wrangling and analysis; replace -9999 values with NA: 

```{r}
fullData <- py_to_r(py$fullData)%>%
  rename(parameterValue = result, dateTime = phenomenon_time, parameter = param)%>% #renaming columns
  mutate(parameterValue = ifelse(parameterValue == -9999, NA, parameterValue))%>% #replace "no data" value o f-9999 with NA
  select(!result_qualifier_codes)%>%
  mutate(dateTime = as.POSIXct(dateTime, tz = "UTC", "%m-%d-%Y %H:%M:%S"))#convert date time to POSIXct

```

##### a. Calculate Wood Camp Spring; add back to dataset

```{r}
wcDat <- fullData%>%
  filter(site == "Logan River at Wood Camp Bridge"|site=="Logan River Above Wood Camp")%>% #extract above and below WC sites
  filter(parameter == "Discharge_cfs")%>% #extract discharge values
  select(dateTime, site, parameter, unit, parameterValue)%>% #drop extra columns
  pivot_wider(id_cols = c(dateTime, parameter, unit), names_from = site, values_from = parameterValue)%>% #pivot to get columns w/ discharge for each site
  rename(aboveWc = 4, wcBridge = 5)%>% #renaming cols
  mutate(woodCampSpring = wcBridge-aboveWc)%>% #calculate spring discharge 
  mutate(wcsCor = ifelse(woodCampSpring<0, NA, woodCampSpring))

#Check data with quick time series:

ggplot(wcDat, aes(x = dateTime, y = woodCampSpring))+
  geom_line()+
  labs(x = "Date", y = "Wood Camp Spring discharge, cfs")+
  theme_classic()

#Plot has a substantial number of negative values (i.e. recorded discharge at WC bridge lower than above WC... WTF?) - checking some of these:

length(which(wcDat$woodCampSpring<0))

wcDat[30915:30925,]
wcDat[31545:31555,]

# No obvious pattern - for now, iteravely replace with average of adjacent values until no more negatives remain (takes a while to run): 

while(length(which(wcDat$woodCampSpring<0))>0){
  wcDat<-wcDat%>%
    mutate(woodCampSpring = ifelse(woodCampSpring<0, #If woodcamp spring is <0
                                 ((lag(woodCampSpring)+lead(woodCampSpring))/2), #replace with average of adjacent rows
                                 woodCampSpring #Else leave as is. 
                                 ))
  
  print(paste(length(which(wcDat$woodCampSpring<0)), "negatives remaining", sep = " "))
}


#check if successful:
length(which(wcDat$woodCampSpring<0)) #Seems to have gotten rid of all negatives. Spot checking lines from above:

wcDat[30915:30925,]
wcDat[31545:31555,]

#Replaced everything, but still results in some very low values. Compare plot of replaced values vs. NA replacement:

(ggplot(wcDat, aes(x = dateTime, y = wcsCor))+
  geom_line()+
  labs(x = "Date", y = "Wood Camp Spring discharge, cfs", title = "Replaced with NA")+
  theme_classic())|(ggplot(wcDat, aes(x = dateTime, y = woodCampSpring))+
  geom_line()+
  labs(x = "Date", y = "Wood Camp Spring discharge, cfs", title = "Filled")+
  theme_classic())

#Use filled data for now - but will have some artificially low values. 

wcFinal <- wcDat%>%
  select(dateTime, parameter, unit, woodCampSpring)%>% #drop extra cols
  mutate(site = rep("Wood Camp Spring", nrow(wcDat)))%>% #add site name
  rename(parameterValue = woodCampSpring) #rename for rbind with maind data

#add back to full dataset: 

fullData_wc <- rbind(fullData, wcFinal)%>%
  filter(!(site == "Logan River at Wood Camp Bridge"|site=="Logan River Above Wood Camp"))
```

#### iv. Filling in missing data for Rick's with 0's

First, identify bounds of when spring is flowing - when are earliest start date, earliest end date? 

```{r}
ricks <- fullData%>%
  filter(site == "Ricks Spring above confluence with Logan River")%>% #extract Ricks data
  mutate(date = ymd(substr(dateTime, 1,10)))%>%
  mutate(doy = yday(dateTime), 
         year = year(dateTime),
         naDoy = ifelse(is.na(parameterValue), NA, doy))%>% #add columns for date, day of year, year, and Na day of year (= NA if data col = NA, otherwise = doy)
  group_by(year, parameter)%>% #group by year and parameter
  summarise(doyMin = min(doy), 
            doyMax = max(doy), #calculate starting and ending day of year for each dataset
            start = first(naDoy, na_rm = T), 
            end = last(naDoy, na_rm = T))%>% #extract first and last obs with non-NA data
  filter(doyMin<121) #drop years that start after May 1st

maxStart <- max(ricks$start)
minEnd <- min(ricks$end) #extract latest start and earliest end DOY's
```

Fill in missing date/times, then replace missing values with 0's:

```{r}
ricksFilled <- fullData%>%
  filter(site == "Ricks Spring above confluence with Logan River")%>%
  group_by(parameter)%>%
  complete(dateTime = seq(min(dateTime), max(dateTime), by = "15 mins"))%>%
  mutate(parameter = ifelse(is.na(parameter), unique(parameter), parameter))%>%
  ungroup()%>%
  mutate(doy =yday(dateTime))%>% #add doy column for calculations
  mutate(parameterValue=ifelse(
    (is.na(parameterValue)&doy<maxStart)|(is.na(parameterValue)&doy>minEnd), #if parameterValue is NA AND is before max start or after min end, 
    0, parameterValue #replace with 0, else leave as is. 
  ))%>%
  mutate(site = ifelse(is.na(site), "Ricks Spring above confluence with Logan River", site))%>%
  mutate(unit = ifelse(parameter == "Discharge_cfs", "cubic feet per second", "degree Celsius"))%>%
  select(!doy)
```

Replace NA values before the latest start date / after the earliest end date with 0's:

```{r}
fullData_ricks <- fullData_wc%>%
  filter(!site == "Ricks Spring above confluence with Logan River")%>% #drop original data
  bind_rows(ricksFilled) #replace with filled data
  

ggplot(filter(fullData_ricks, site =="Ricks Spring above confluence with Logan River"), aes(x = dateTime, y = parameterValue, color = parameter))+
  geom_point()+
  theme_classic()
```


#### v. Adding Logan River @ First Dam and Highline Canal

Read in data from USGS - NOTE current USGS API function only supports calling up to 3yrs of data at a time, so using a for-loop to pull and concatenate all data back to 1986 - takes a while to run. 

```{r}
sitesUSGS <- c("USGS-10109000", "USGS-10108400") #USGS site ID's for Logan River at 1st Dam and Highline Canal
USGSparam <- "00060"#parameter code for discharge
dataPeriod <- c(seq.Date(from = "1986-10-01", to = Sys.Date(), by = "year"), NA) #make vector of start/end dates to use in loop
USGSdat <- data.frame() #set up empty dataframe to recieve data

#read in data with for loop: 

for(i in 1:length(dataPeriod)){
  d <- read_waterdata_continuous(monitoring_location_id = sitesUSGS, 
                                        parameter_code = USGSparam, #target sites and parameter identified above
                                       time = c(dataPeriod[i], dataPeriod[(i+1)]) #setting data period - using 1-year increments
                                       )
  USGSdat <- rbind(USGSdat, d) #rbind i'th year's data to USGSdat dataframe
  
  print(paste("finished", year(dataPeriod[i]), sep = " ")) #print year finished for tracking. 
}

#check for success:

head(USGSdat)
tail(USGSdat)
```

Next, drop unneeded columns and calculate total discharge for first dam (first dam + highline canal) for each date/time. 

```{r}
USGSpivot <- USGSdat%>%
  select(monitoring_location_id, time, value)%>% #extract site, time, and value columns
  distinct()%>% #drop duplicate rows
  pivot_wider(id_cols = time, names_from = monitoring_location_id, values_from = value)%>% #pivot wider to get rows for each dateTime, columns with values for each site
  mutate(`USGS-10108400` = ifelse(is.na(`USGS-10108400`), 0, `USGS-10108400`)) #replace NA values in Highline with 0's


loganAtFirstDam <- USGSpivot%>%
  mutate(parameterValue = `USGS-10108400`+`USGS-10109000`, 
         site = rep("Logan River at First Dam", nrow(USGSpivot)), 
         parameter = rep("Discharge_cfs", nrow(USGSpivot)), 
         unit = rep("cubic feet per second", nrow(USGSpivot)))%>%
  rename(dateTime = time)%>%
  select(!c(`USGS-10108400`, `USGS-10109000`))
```

NOTE - switched from hourly readings to readings every 15min - check when switch was made:

```{r}
obsPerYear <- loganAtFirstDam%>%
  mutate(year = year(dateTime))%>%
  group_by(year)%>%
  summarize(nObs = n())

View(obsPerYear)
```

First year with 15min intervals is **1995**. All data from 1989-1994 is in 1hr intervals. 

Add Logan at First Dam to overall dataset: 

```{r}
fullData_usgs <- rbind(fullData_ricks, loganAtFirstDam)
```

Check date ranges for each site: 

```{r}
dateRanges <- fullData_usgs%>%
  group_by(site, parameter)%>%
  summarise(start = min(dateTime, na.rm = T), 
            end = max(dateTime, na.rm = T))%>%
  mutate(minDoy = yday(start), 
         maxDoy = yday(end))
dateRanges
```


#### vi. Data Cleaning + QA/QC

First QA/QC Step: identifying partial years; create second dataset with only full years of data

```{r}
fullData_partial <- fullData_usgs%>%
  mutate(year=year(dateTime), 
         doy = yday(dateTime))%>% #add columns for year and day of year
  group_by(site, parameter, year)%>%
  mutate(doyMax = max(doy, na.rm = T), 
         doyMin = min(doy, na.rm = T))%>% #calculate minimum and maxiumum day of year for each site + parameter
  mutate(partialYear = ifelse((doyMax<365|doyMin>1), T, F))%>% #add T/F column flagging years with partial data (start day of year >1 or end day of year <365)
  ungroup() #ungroup

fullData_complete <- fullData_partial %>%
  filter(!partialYear == T) #create new dataset with only complete years of data
```

Check full years (i.e. data for 1/1 and 12/31) for data gaps:

```{r}
dataGaps <- fullData_complete%>%
  mutate(date = ymd(substr(dateTime, 1,10)))%>% #create new column with date
  group_by(site, parameter, date, year)%>% #group by site/parameter/date
  summarise(
    dailyMean = mean(parameterValue) #calculate mean parameter value for each site/parameter/day - will set any day that includes NA data to NA
  )%>%
  mutate(dataPresent = ifelse(is.na(dailyMean), 0, 1))%>% #re-code days with NA data as 0, days with data as 1
  group_by(site, parameter, year)%>%
  summarise(dataDays = sum(dataPresent))%>% #group by year and calculate total # days with data for each site/parameter
  mutate(missingDays = ifelse(dataDays == 366, (366-dataDays), (365-dataDays)), 
         siteParam = paste(site, parameter, sep = " ")) #calculate # missing days, accounting for leap years; add site+parameter column
  
#View(dataGaps)
```


Data gap visualizations:

```{r}
dg.plot <- ggplot(dataGaps, aes(x = year, y = dataDays, fill = parameter))+
  geom_bar(stat = "identity", position = "dodge")+
  geom_hline(aes(yintercept = 365), colour = "darkgrey", linetype = "dashed")+ #add grey dashed line @ y = 365 days
  geom_hline(aes(yintercept = 265), colour = "darkred", linetype = "dashed")+ #add red dashed line @ y = 265 days (>100 days missing)
  facet_wrap(~site, scales = "free_x")+
  labs(x = "Year", y = "# days with data recorded", fill = "Parameter")+
  scale_fill_manual(values = c("cyan4", "darkblue", "cyan2", "lightblue"))+
  theme_classic()+
  theme(strip.text = element_text(size = 10, face = "bold"), 
        axis.title = element_text(size = 10, face = "bold"), 
        axis.text = element_text(size = 10))
dg.plot

#Save output: 

ggsave("DataGaps.jpg", dg.plot, device = "jpeg", path = here("outputs", "plots"), units = "in", dpi = "retina", height = 6, width = 11)

```

Many years with >100 days of missing data for discharge at Franklin Basin and Rick's Spring, most other sites ok. Flag years with <100 missing days:

```{r}
dataGaps <- dataGaps%>%
  mutate(flag = ifelse(missingDays>100, T, F)) #add flag column to mark years with >100 missing days
```

Add flags and missing days back into full dataset

```{r}
fullData_flagged <- merge(dataGaps,fullData_complete)
```


#### vii. Alternative formatting for DOY plotting:

Alternative data format for DOY plots: 

```{r}
dailyData <- fullData_usgs%>%
  mutate(parameter = ifelse(parameter == "Discharge_cfs", "Discharge", parameter),
    siteParam = paste(site, parameter, sep = " "))%>% #add new Site + parameter column for grouping
  mutate(doy = yday(dateTime), #make new DOY column
         date = ymd(substr(dateTime, 1,10)))%>% #make new column with just date
  group_by(siteParam, date, unit, parameter, doy)%>%
  summarise(paramDaily = mean(parameterValue, na.rm = TRUE))%>% #calculate daily average
  group_by(siteParam, doy)%>% #group by day of year
  mutate(paramMin = min(paramDaily, na.rm = T), 
         paramMax = max(paramDaily, na.rm = T), 
         paramMed = median(paramDaily, na.rm = T), 
         paramMean = mean(paramDaily, na.rm = T), 
         qt1 = quantile(paramDaily, probs = 0.1, na.rm = T)[[1]], 
         qt2 = quantile(paramDaily, probs = 0.3, na.rm = T)[[1]], 
         qt3 = quantile(paramDaily, probs = 0.7, na.rm = T)[[1]], 
         qt4 = quantile(paramDaily, probs = 0.9, na.rm = T)[[1]], 
         year = year(date))%>% #calculate day of year minimum, maximum, median, and mean, plus quantiles for each doy
  ungroup()%>% #ungroup
  arrange(siteParam, date) #sort by site+parameter and date for easier checking


unique(dailyData$siteParam)
```

### B. SNOTEL Data

Import data from Snotel using the snotelr package:

Get metadata for sites in the Logan River Watershed:

```{r}
snotelSites <- snotel_info(network="sntl")%>%
  filter(str_detect(description, "Logan")|str_detect(description, "Temple")) #filter out sites in Logan River Watershed

lro_siteIds <- c(snotelSites$site_id)

snotelAll <- snotel_info(network="sntl")%>%
  filter(county == "Cache"|county=="Rich"|county=="Franklin"|county=="Bear Lake") #Info for sites in Cache, Rich, Franklin, and Bear Lake counties

all_siteIds <- c(snotelAll$site_id)
```

NOTE - currently NOT including USU Doc Daniel (on drainage divide between Logan and Garden City sides) or Little Bear (S End of Cache Valley)

Download data for all sites in the logan watershed: 

```{r}
loganSnotel <- snotel_download(site_id = lro_siteIds, network = "sntl", internal = T)%>%
  mutate(date = ymd(date), site_name= as.factor(site_name))
```

Data for all sites in the logan watershed + surrounding counties:

```{r}
allSnotel <- snotel_download(site_id = all_siteIds, network = "sntl", internal = T)%>%
  mutate(date = ymd(date), site_name = as.factor(site_name))
```

## II. Basic Time Series Plots

### A. Discharge

Starting with a basic time series, y = discharge CFS, x = date Time, color = site. 

```{r}
#color palate: 

discharge.pal <- brewer.pal(length(unique(fullData_usgs$site)), "BrBG")


#Check with updated plot: 
q.comb <- ggplot(
  filter(fullData_usgs, parameter == "Discharge_cfs"), #extract discharge data
  aes(x = dateTime, y = parameterValue, color = site) #setting mapping
)+
  geom_line()+
  theme_classic()+
  labs(x = "Date", y = "Discharge, CFS", color = "Station Name")+ #axis labels
  scale_color_manual(values = discharge.pal)+ #Setting color palate
  theme(legend.position = "bottom", 
        axis.title = element_text(size = 10, face = "bold"), 
        legend.title = element_text(size = 10, face = "bold")) #editing text appearance 

q.comb

#Save: 

ggsave("Discharge_Comb.jpg", q.comb, device = "jpeg", path = here("outputs", "plots"), units = "in", dpi = "retina", height = 6, width = 11)
```

### B. Water Temperature

Basic Time Series:

```{r}
#Initial plot:
stream.temp <- ggplot(
  filter(fullData,
         parameter == "Water Temperature"), 
  aes(x = dateTime, y = parameterValue, color = site)
)+
  geom_line()+
  theme_classic()+
  labs(x = "Date", y = "Stream Temperature, deg. C", color = "Station Name")+
  scale_color_manual(values = discharge.pal)+
  theme(legend.position = "bottom", 
        axis.title = element_text(size = 10, face = "bold"), 
        legend.title = element_text(size = 10, face = "bold"))

stream.temp

#Save: 

ggsave("StreamTemp_Comb.jpg", stream.temp, device = "jpeg", path = here("outputs", "plots"), units = "in", dpi = "retina", height = 6, width = 11)
```



### C. Air Temperature

Basic time series:

```{r}
climate.pal <- c("#C7E9C0","#238B45" )

air.temp <- ggplot(
  filter(fullData,
         parameter == "Air Temperature"), 
  aes(x = dateTime, y = parameterValue, color = site)
)+
  geom_line()+
  theme_classic()+
  labs(x = "Date", y = "Air, deg. C", color = "Station Name")+
  scale_color_manual(values = climate.pal)+
  theme(legend.position = "bottom", 
        axis.title = element_text(size = 10, face = "bold"), 
        legend.title = element_text(size = 10, face = "bold"))

air.temp


#Save: 

ggsave("AirTemp_Comb.jpg", air.temp, device = "jpeg", path = here("outputs", "plots"), units = "in", dpi = "retina", height = 6, width = 11)
```

### D. Precipitation

Basic time series:

```{r}
precip <- ggplot(
  filter(fullData,
         parameter == "Precipitation"), 
  aes(x = dateTime, y = parameterValue, color = site)
)+
  geom_line()+
  theme_classic()+
  labs(x = "Date", y = "Cumulative Annual Precipitation, cm", color = "Station Name")+
  scale_color_manual(values = climate.pal)+
  theme(legend.position = "bottom", 
        axis.title = element_text(size = 10, face = "bold"), 
        legend.title = element_text(size = 10, face = "bold"))

precip


#Save: 

ggsave("Precip_Comb.jpg", precip, device = "jpeg", path = here("outputs", "plots"), units = "in", dpi = "retina", height = 6, width = 11)
```

### E. SWE

```{r}
swe <- ggplot(loganSnotel, aes(x = date, y = snow_water_equivalent, color = site_name))+
  geom_line()+
  theme_classic()+
  labs(x = "Date", y = "Snow Water Equivalent, cm", color = "Site Name")+
  scale_color_manual(values = c("darkblue", "blue", "lightblue", "cyan4"))+
  facet_wrap(~site_name)+
  theme(legend.position = "none", 
        axis.title = element_text(size = 10, face = "bold"), 
        legend.title = element_text(size = 10, face = "bold"), 
        strip.text = element_text(size = 10, face = "bold"))
swe

#Save: 

ggsave("SWE_Comb.jpg", swe, device = "jpeg", path = here("outputs", "plots"), units = "in", dpi = "retina", height = 6, width = 11)

```


## III. Day of year plotting

Using "dailyData" dataframe created in step I. 

### A. Testing

First, Test plotting day of year plot with one site - data prep: 

```{r}
thisyear <- format(Sys.Date(), "%Y") #Create object with this year for later filtering
test <- filter(dailyData, siteParam == "Climate Station at Franklin Basin Air Temperature")%>%
  filter(!year<(max(year)-5))#make test dataset with one site/parameter

datemin <- year(min(test$date))
datemax <- year(max(test$date)) #Make date range object for labeling
```

Plotting:

```{r}
testplot <- ggplot()+ 
  geom_line(data = test, aes(x = doy, y = paramMin, color = "Minimum"))+ #draw minimum line 
  geom_line(data = test, aes(x = doy, y = paramMed, color = paste("Median ", "(", datemin, "-", datemax, ")", sep = "")))+ #draw median line, add daterange object for label
  geom_line(data = test, aes(x = doy, y = paramMax, color = "Maximum"))+ #add maximum line
  geom_line(data = filter(test, year(date) == datemax), aes(x = doy, y = paramDaily, color = paste(datemax)))+ #filter out this year's data and add to plot
  scale_color_manual(values = c("black", "blue", "green", "red"))+ #Setting line color scale
  labs(x = "Day of Year", y = "Discharge, CFS", color = "Legend", title = unique(test$siteParam))+ #axis and legend lables
  geom_ribbon(data = test, aes(x = doy, ymax = paramMax, ymin = qt4), fill = "blue", alpha = 0.25)+
  geom_ribbon(data = test, aes(x = doy, ymax = qt4, ymin = qt3), fill = "lightblue", alpha = 0.25)+
  geom_ribbon(data = test, aes(x = doy, ymax = qt3, ymin = qt2), fill = "green", alpha = 0.25)+
  geom_ribbon(data = test, aes(x = doy, ymax = qt2, ymin = qt1), fill = "yellow", alpha = 0.25)+
  geom_ribbon(data = test, aes(x = doy, ymax = qt1, ymin = paramMin), fill = "red", alpha = 0.25)+ #adding shaded ribbons for quantiles
  theme_classic()+ #removing background formatting
  theme(legend.position = c(0.8, 0.75), #repositioning legend within plot area
        axis.text = element_text(size = 10), 
        axis.title = element_text(size = 10, face = "bold"), 
        legend.text = element_text(size = 9), 
        legend.title = element_text(size = 9, face = "bold"), 
        plot.title = element_text(size = 11, face = "bold")) #axis and legend text formatting
testplot
```

Success! 

### B. Make DOY Plots w/ For-Loop

Next, iterate over each site/parameter to make all plots: 

Setup: 

```{r}
allSiteParams <- unique(dailyData$siteParam) #make vector of all site/parameter combos
plot.list <- list() #empty list to recieve plots
thisyear <- format(Sys.Date(), "%Y") #Create object with this year for later filtering
q.pal <- c("#0000FF35", "#ADD8E635", "#00FF0035", "#FFFF0035" ,"#FF000035")
t.pal <- rev(q.pal)
q.line <- c("black", "blue", "green", "red")
t.line <- c("black", "red", "green", "blue") #set up color palate options

dateMaxes <- year(dateRanges$end)
dateMins <- dateMaxes-5
  
```

Make plots with a for-loop:

```{r}
for(i in 1:length(allSiteParams)){
  d <- filter(dailyData, siteParam == allSiteParams[i])%>% #Filter data to "ith" site/parameter
    filter(!year<(max(year)-5))

  fill.pal <- if (str_detect(allSiteParams[[i]], "Temperature")) q.pal else t.pal
  line.pal <- if (str_detect(allSiteParams[[i]], "Temperature")) t.line else q.line
  
  yLab <- paste(
    unique(d$parameter), 
    unique(d$unit), 
    sep = ", "
  )
  
  #Plotting:
  
  p <- ggplot()+ 
  geom_line(data = d, aes(x = doy, y = paramMin, color = "Minimum"))+ #draw minimum line 
  geom_line(data = d, aes(x = doy, y = paramMed, color = paste("Median ", "(", dateMins[[i]], "-", dateMaxes[[i]], ")", sep = "")))+ #draw median line, add daterange object for label
  geom_line(data = d, aes(x = doy, y = paramMax, color = "Maximum"))+ #add maximum line
  scale_color_manual(values = line.pal)+ #Setting line color scale
  labs(x = "Day of Year", y = yLab, color = "Legend", title = unique(d$siteParam), fill = "Percentiles")+ #axis and legend lables
  geom_ribbon(data = d, aes(x = doy, ymax = paramMax, ymin = qt4, fill = "90-100th"))+
  geom_ribbon(data = d, aes(x = doy, ymax = qt4, ymin = qt3, fill = "70-90th"))+
  geom_ribbon(data = d, aes(x = doy, ymax = qt3, ymin = qt2, fill = "30-70th"))+
  geom_ribbon(data = d, aes(x = doy, ymax = qt2, ymin = qt1, fill = "10-30th"))+
  geom_ribbon(data = d, aes(x = doy, ymax = qt1, ymin = paramMin, fill = "0-10th"))+ #adding shaded ribbons for quantiles
  scale_fill_manual(values = fill.pal)+
  geom_line(data = filter(d, year(date) == dateMaxes[[i]]), aes(x = doy, y = paramDaily, color = as.character(dateMaxes[[i]])), linewidth = 0.75)+ #filter out this year's data and add to plot
    #guides(fill = "none")+
  theme_classic()+ #removing background formatting
  theme(legend.position = "bottom", #re positioning legend along bottom
        axis.text = element_text(size = 11), 
        axis.title = element_text(size = 11, face = "bold"), 
        legend.text = element_text(size = 10), 
        legend.title = element_text(size = 10, face = "bold"), 
        plot.title = element_text(size = 12, face = "bold"),
         #legend.key = element_rect(fill = alpha(fill.pal, 0.25))
        ) #axis and legend text formatting
    
  
  plot.list[[i]]<-p #Save plot as "ith" plot in plot list
  
  fn <- paste(allSiteParams[[i]], ".jpg", sep = "")
  
  ggsave(fn, p, device = "jpeg", path = here("outputs", "plots", "doy_plots"), units = "in", dpi = "retina", height = 6, width = 11)
  
  print(paste("Finished", allSiteParams[[i]], sep = " ")) #Print site name/parameter for tracking
}
```

Check outputs: 

```{r}
plot.list[[7]]
```

## IV. Basic Stats

### A. Annual Trend Analyses

Data Prep: 

```{r}
yearData <- dailyData%>%
  mutate(year = year(date))%>% #create year column
  group_by(siteParam, year, parameter, unit)%>% #group by site/parameter and year
  summarise(annualTot = sum(paramDaily, na.rm = T), 
            annualMax = max(paramDaily, na.rm = T), 
            annualMin = min(paramDaily, na.rm = T), 
            annualMean = mean(paramDaily, na.rm = T), 
            annualMed = median(paramDaily, na.rm = T), #calculate annual total, max, min, mean, and median for each site/param
            maxDate = date[which.max(paramDaily)], 
            minDate = date[which.min(paramDaily)])%>% #extract date for maximum and minimum values
  mutate(annualTot = ifelse(parameter == "Precipitation", annualMax, annualTot)) #correcting precip data (already cumulative)
```

Basic LM for annual total x year:

First, nest data based on site: 

```{r}
yearNested <- yearData %>%
  nest(data = -siteParam) #create new nested dataframe with one column for site, one column for data - data entries for each site include year and mean temp. 
```

Next, run a model for each site separately and return a summary table with the results for each site:

```{r}
yearModels <- yearNested %>%
  mutate(model = map(data, ~lm(annualTot~year, data = .)%>% #run a model of temp~year for each site (referenced by "."); store results in new column called "model"
                     tidy))%>% #tidy function to construct a summary table with model results (seperate tables for each site)
  unnest(model) %>% #unnest model column to get rows for each site, columns for model metrics
  filter(term == "year")  #remove intercept term to just get info on year for each site
yearModels
```



NOT RUN: Plotting - annual total: 


annual.plots <- list()


for(i in 1:length(allSiteParams)){
  df <- filter(yearData, siteParam == allSiteParams[i])
  
  pv <- yearModels$p.value[i]
  
  y.l <- paste("Total annual", unique(df$parameter), "," ,unique(df$unit), sep = " ")
  
  pl <- ggplot(df, aes(x = year, y = annualTot))+
    geom_point()+
    geom_smooth(se = T, method = "lm", linetype = "dashed", color = "darkgrey")+
    theme_classic()+
    labs(x = "Year", y = y.l, title = allSiteParams[[i]])+
    geom_text(aes(
      y = (0.9*max(df$annualTot, na.rm = T)), 
      x = (0.9*max(df$year, na.rm = T)),
      label = paste("P=", round(pv, 3), sep = "")))
  
  annual.plots[[i]]<-pl
  
  print(paste("Finished", allSiteParams[[i]]))
  
}

Data prep for facet plot:

```{r}
#Data prep: 

yearPlot <- merge(yearData, yearModels)

pVals <- yearPlot%>%
  group_by(siteParam)%>%
  summarize(x = mean(year), label = round(unique(p.value), 2))

climateStations <- filter(yearPlot, str_detect(siteParam, "Climate"))
streamStations <- filter(yearPlot, !str_detect(siteParam, "Climate"))

```

Plotting for climate stations:

```{r}
#Plotting - Climate Stations:

climate.plot <- ggplot(climateStations, aes(x = year, y = annualTot))+
  geom_point()+
  geom_smooth(se = T, method = "lm", linetype = "dashed", color = "darkgrey")+
  theme_classic()+
  labs(x = "Year", y = "Total annual value")+
  geom_text(data = filter(pVals, str_detect(siteParam, "Climate")), aes(x = x, label = paste("P=", label)), y = Inf, vjust = 2)+
  facet_wrap(~siteParam, scales = "free")+
  theme(axis.text = element_text(size = 11), 
        axis.title = element_text(size = 11, face = "bold"), 
        strip.text = element_text(size = 10, face = "bold"))


climate.plot

#Save plot: 

ggsave("Annual_Climate_Totals.jpg", climate.plot, device = "jpeg", path = here("outputs", "plots"), units = "in", dpi = "retina", height = 6, width = 11)
```

Plotting for stream stations: 

```{r}

stream.plot <- ggplot(streamStations, aes(x = year, y = annualTot))+
  geom_point()+
  geom_smooth(se = T, method = "lm", linetype = "dashed", color = "darkgrey")+
  theme_classic()+
  labs(x = "Year", y = "Total annual value")+
  geom_text(data = filter(pVals, !str_detect(siteParam, "Climate")), aes(x = x, label = paste("P=", label)), y = Inf, vjust = 2)+
  facet_wrap(~siteParam, scales = "free")+
  theme(axis.text = element_text(size = 11), 
        axis.title = element_text(size = 11, face = "bold"), 
        strip.text = element_text(size = 10, face = "bold"))


stream.plot

#Save plot:
ggsave("Annual_Stream_Totals.jpg", stream.plot, device = "jpeg", path = here("outputs", "plots"), units = "in", dpi = "retina", height = 11, width = 12)
```

### B. Individual Variable Analyses:

For each primary variable, measures of interest include: 

- **Discharge:**
  + Total annual discharge
  + Timing of peak discharge
  + Mean annual discharge
- **Stream Temperature:**
  + Annual average stream temperature
  + "Degree days" over a certain threshold - maybe 20C?
  + Average Late Summer Stream Temp (July-Aug?)
- **Precip:**
  + Total annual precip
- **Air Temperature**
  + Average annual air temperature
  + "Degree days" over a certain threshold?
  + Average seasonal temps: Winter (December-February), Spring (March-May), Summer(June-August), Fall (September-November)
- **SWE:**
  + Peak SWE
  + Date of peak SWE
  + "Meltout time": # days between peak SWE and zero SWE
  
#### i. Discharge:

First, extract data and calculate total annual discharge, mean annual discharge, and timing of peak discharge; reorganize for analysis and plotting:

```{r}
discharge <- yearData%>%
  filter(parameter == "Discharge")%>% #extract discharge data from prev. calculated annual min, mean, max, etc. 
  mutate(maxDOY =yday(maxDate))%>% #add day of year for max discharge
  select(siteParam, year, parameter, unit, annualTot, annualMean, maxDOY)%>%
  pivot_longer(!c(siteParam, year, parameter, unit), names_to = "metricName", values_to = "metricValue")%>%
  filter(!is.na(year))
```

Nest data for modeling:  

```{r}
dischargeNest <- discharge %>%
  nest(data = -c(siteParam, metricName)) #create new nested dataframe with columns for site and metricName, one column for data - data entries for each site include year and metricValues. 
```

Run basic models for each site and metric of interest: 

```{r}
qModels <- dischargeNest %>%
  mutate(model = map(data, ~lm(metricValue~year, data = .)%>% #run a model of temp~year for each site (referenced by "."); store results in new column called "model"
                     tidy))%>% #tidy function to construct a summary table with model results (seperate tables for each site)
  unnest(model) %>% #unnest model column to get rows for each site, columns for model metrics
  filter(term == "year")  #remove intercept term to just get info on year for each site
qModels
```

Add P-values for plotting: 

```{r}
dischargePV <- merge(discharge, 
                     select(qModels,siteParam, metricName, p.value)
                     )%>%
  mutate(metricPrint = case_when(
    metricName == "annualMean"~ "Mean Annual Discharge, cfs", 
    metricName == "annualTot" ~"Total Annual Discharge, cfs", 
    metricName == "maxDOY" ~"Max. Discharge Day of Year"
  ))%>%
  mutate(site = str_remove(siteParam, "Discharge"), 
         siteMetric = paste(site, metricPrint, sep = "- "))

qPVs <- dischargePV%>%
  group_by(siteMetric, metricPrint)%>%
  summarize(x = mean(year), label = round(unique(p.value), 2))
```

Draw plots: 

```{r}
q.plot <- ggplot(dischargePV, aes(x = year, y = metricValue, color = metricPrint))+
  geom_point()+
  geom_smooth(se = T, method = "lm", linetype = "dashed")+
  theme_classic()+
  labs(x = "Year", y = "Metric Value", color = "Metric")+
  scale_color_manual(values = c("cyan4", "darkblue", "darkgreen"))+
  geom_text(data = qPVs, aes(x = x, label = paste("P=", label)), y = Inf, vjust = 2)+
  facet_wrap(~siteMetric, scales = "free", labeller = label_wrap_gen(width = 40))+
  theme(axis.text = element_text(size = 11), 
        axis.title = element_text(size = 11, face = "bold"), 
        strip.text = element_text(size = 10, face = "bold"), 
        legend.position = "bottom")


q.plot

#Save plot:
ggsave("Discharge_metrics.jpg", q.plot, device = "jpeg", path = here("outputs", "plots"), units = "in", dpi = "retina", height = 10, width = 10)
```

#### ii. Stream Temperature

First, extract data and calculate annual average stream temperature, # days with average temp >20C, and average July-Aug stream temp; reorganize for analysis and plotting:

```{r}
streamTemp <- dailyData%>%
  filter(parameter == "Water Temperature")%>% #extract discharge data from prev. calculated annual min, mean, max, etc. 
  mutate(month = month(date),
         year = year(date), 
         summerT = ifelse(month == 7|month ==8, paramDaily, NA), 
         over20 = ifelse(paramDaily >20, 1, 0))%>%
  group_by(siteParam, year, parameter, unit)%>%
  summarize(stMean = mean(paramDaily, na.rm =T), 
            stSummer = mean(summerT, na.rm = T), 
            stDegDay = sum(over20, na.rm = T)
            )%>%
  pivot_longer(!c(siteParam, year, parameter, unit), names_to = "metricName", values_to = "metricValue")%>%
  filter(!is.na(year))
```

Nest data for modeling:  

```{r}
stNest <- streamTemp %>%
  nest(data = -c(siteParam, metricName)) #create new nested dataframe with columns for site and metricName, one column for data - data entries for each site include year and metricValues. 
```

Run basic models for each site and metric of interest: 

```{r}
stModels <- stNest %>%
  mutate(model = map(data, ~lm(metricValue~year, data = .)%>% #run a model of temp~year for each site (referenced by "."); store results in new column called "model"
                     tidy))%>% #tidy function to construct a summary table with model results (seperate tables for each site)
  unnest(model) %>% #unnest model column to get rows for each site, columns for model metrics
  filter(term == "year")  #remove intercept term to just get info on year for each site
stModels
```

Add P-values for plotting: 

```{r}
streamTempPV <- merge(streamTemp, 
                     select(stModels,siteParam, metricName, p.value)
                     )%>%
  mutate(metricPrint = case_when(
    metricName == "stMean"~ "Mean Annual Stream Temperature, C", 
    metricName == "stSummer" ~"Mean July-Aug Stream Temperature, C", 
    metricName == "stDegDay" ~"# days with mean stream temperature >20 C"
  ))%>%
  mutate(site = str_remove(siteParam, "Surface Water Temperature"), 
         siteMetric = paste(site, metricPrint, sep = "- "))

stPVs <- streamTempPV%>%
  group_by(siteMetric, metricPrint)%>%
  summarize(x = mean(year), label = round(unique(p.value), 2))
```

Draw plots: 

```{r}
st.plot <- ggplot(streamTempPV, aes(x = year, y = metricValue, color = metricPrint))+
  geom_point()+
  geom_smooth(se = T, method = "lm", linetype = "dashed")+
  theme_classic()+
  labs(x = "Year", y = "Metric Value", color = "Metric")+
  scale_color_manual(values = c("cyan4", "darkblue", "darkgreen"))+
  geom_text(data = stPVs, aes(x = x, label = paste("P=", label)), y = Inf, vjust = 2)+
  facet_wrap(~siteMetric, scales = "free", labeller = label_wrap_gen(width = 40))+
  theme(axis.text = element_text(size = 11), 
        axis.title = element_text(size = 11, face = "bold"), 
        strip.text = element_text(size = 10, face = "bold"), 
        legend.position = "bottom")


st.plot

#Save plot:
ggsave("StreamTemp_metrics.jpg", st.plot, device = "jpeg", path = here("outputs", "plots"), units = "in", dpi = "retina", height = 10, width = 10)
```

#### iii. Precip:
  + Total annual precip
  + Total annual precip on days with temp <32F (proxy for snow vs. rain)
  
First, extract precip data and calculate total annual precip, total precip on days with temp <32F (proxy for snow). 

```{r}
sites <- fullData%>%
  mutate(siteParam = paste(site, parameter, sep = " "))%>%
  select(site, siteParam)%>%
  unique() #extract site names


precip <- merge(dailyData, sites)%>% #add site names 
  filter(parameter == "Air Temperature"|parameter == "Precipitation")%>% #extract precip and air temp data
  pivot_wider(id_cols = c(site, date), names_from = parameter, values_from = paramDaily)%>% #get cols for each parameter
  mutate(year = year(date))%>%
  rename(airTemp = 3)%>%
  arrange(site, date)%>%
  mutate(siteYr = paste(site, year, sep = " ")) #add year column, re-arrange for calculations, add site+year column
  
precipSplit <- split(precip, precip$siteYr) #split into list of DF's for each site+year combo

precipLag <- list() #empty list to recieve new values

for(i in 1:length(precipSplit)){
  x <- precipSplit[[i]]%>%
    mutate(dailyPrecip = (Precipitation-lag(Precipitation)))
  
  precipLag[[i]]<-x #use for loop to calculate daily precip (current day precip minus previous day precip) for each site + year combo
}

precipComb <- bind_rows(precipLag)%>%
  mutate(dailyPrecip = ifelse(dailyPrecip<0, 0, dailyPrecip))%>% #set negative daily precip values to zero
  mutate(coldPrecip = ifelse(airTemp <0, dailyPrecip, 0))%>% #cold precip column - 0 precip if air temp >0, otherwise precip column value
  group_by(site, year)%>%
  summarise(totPrecip = max(Precipitation, na.rm = T), 
            totCold = sum(coldPrecip, na.rm = T))%>% #calculate total annual precip + total cold precip for each site+year combo
  mutate(totCold = ifelse(totCold == 0, NA, totCold))%>% #set 0 tot cold values to NA (no air temp values)
  pivot_longer(!c(site, year), names_to = "metricName", values_to = "metricValue")%>% #pivot longer for analysis and plotting. 
  filter(!is.na(year))
```

Nest data for modeling:  

```{r}
precipNest <- precipComb %>%
  nest(data = -c(site, metricName)) #create new nested dataframe with columns for site and metricName, one column for data - data entries for each site include year and metricValues. 
```

Run basic models for each site and metric of interest: 

```{r}
pcModels <- precipNest %>%
  mutate(model = map(data, ~lm(metricValue~year, data = .)%>% #run a model of temp~year for each site (referenced by "."); store results in new column called "model"
                     tidy))%>% #tidy function to construct a summary table with model results (seperate tables for each site)
  unnest(model) %>% #unnest model column to get rows for each site, columns for model metrics
  filter(term == "year")  #remove intercept term to just get info on year for each site
pcModels
```

Add P-values for plotting: 

```{r}
precipPV <- merge(precipComb, 
                     select(pcModels,site, metricName, p.value)
                     )%>%
  mutate(metricPrint = case_when(
    metricName == "totPrecip"~ "Total Annual Precipitation, cm", 
    metricName == "totCold" ~"Total Annual Precip. on Days <0C, cm"))%>%
  mutate(siteMetric = paste(site, metricPrint, sep = "- "))

pcPVs <- precipPV%>%
  group_by(siteMetric, metricPrint)%>%
  summarize(x = mean(year), label = round(unique(p.value), 2))
```

Draw plots: 

```{r}
pc.plot <- ggplot(precipPV, aes(x = year, y = metricValue, color = metricPrint))+
  geom_point()+
  geom_smooth(se = T, method = "lm", linetype = "dashed")+
  theme_classic()+
  labs(x = "Year", y = "Metric Value", color = "Metric")+
  scale_color_manual(values = c("cyan4", "darkblue"))+
  geom_text(data = pcPVs, aes(x = x, label = paste("P=", label)), y = Inf, vjust = 2)+
  facet_wrap(~siteMetric, scales = "free", labeller = label_wrap_gen(width = 40))+
  theme(axis.text = element_text(size = 11), 
        axis.title = element_text(size = 11, face = "bold"), 
        strip.text = element_text(size = 10, face = "bold"), 
        legend.position = "bottom")


pc.plot

#Save plot:
ggsave("Precip_metrics.jpg", pc.plot, device = "jpeg", path = here("outputs", "plots"), units = "in", dpi = "retina", height = 10, width = 10)
```


#### iv. Air Temperature 
  + Average annual air temperature
  + "Degree days" over a certain threshold?
  + Average seasonal temps: Winter (December-February), Spring (March-May), Summer(June-August), Fall (September-November)

**NOTE - Currently have winter temperature as Jan-Feb + December of same CALENDAR year, preferable to have December (year 1)- Feb (year 2).** 

First, extract data and calculate annual average stream temperature, # days with average temp >20C, and average July-Aug stream temp; reorganize for analysis and plotting:

```{r}
airTemp <- dailyData%>%
  filter(parameter == "Air Temperature")%>% #extract discharge data from prev. calculated annual min, mean, max, etc. 
  mutate(month = month(date),
         year = year(date), 
         winterT = ifelse(month == 1|month ==2| month == 12, paramDaily, NA),
         springT = ifelse(month == 3|month == 4| month == 5, paramDaily, NA),
         summerT = ifelse(month ==6|month == 7|month ==8, paramDaily, NA),
         fallT = ifelse(month == 9|month == 10|month == 11, paramDaily, NA),
         over25 = ifelse(paramDaily >25, 1, 0))%>%
  group_by(siteParam, year, parameter, unit)%>%
  summarize(atMean = mean(paramDaily, na.rm =T), 
            atWinter = mean(winterT, na.rm =T), 
            atSpring = mean(springT, na.rm = T), 
            atSummer = mean(summerT, na.rm = T), 
            atFall = mean(fallT, na.rm = T),
            atDegDay = sum(over25, na.rm = T)
            )%>%
  pivot_longer(!c(siteParam, year, parameter, unit), names_to = "metricName", values_to = "metricValue")%>%
  filter(!is.na(year))
```

Nest data for modeling:  

```{r}
atNest <- airTemp %>%
  nest(data = -c(siteParam, metricName)) #create new nested dataframe with columns for site and metricName, one column for data - data entries for each site include year and metricValues. 
```

Run basic models for each site and metric of interest: 

```{r}
atModels <- atNest %>%
  mutate(model = map(data, ~lm(metricValue~year, data = .)%>% #run a model of temp~year for each site (referenced by "."); store results in new column called "model"
                     tidy))%>% #tidy function to construct a summary table with model results (seperate tables for each site)
  unnest(model) %>% #unnest model column to get rows for each site, columns for model metrics
  filter(term == "year")  #remove intercept term to just get info on year for each site
atModels
```

Add P-values for plotting: 

```{r}
airTempPV <- merge(airTemp, 
                     select(atModels,siteParam, metricName, p.value)
                     )%>%
  mutate(metricPrint = case_when(
    metricName == "atMean"~ "Mean Annual Air Temperature, C", 
    metricName == "atWinter"~"Mean Jan/Feb/Dec Air Temperature, C",
    metricName == "atSpring"~"Mean Mar-May Air Temperature, C", 
    metricName == "atFall"~"Mean Sep-Nov Air Temperature, C",
    metricName == "atSummer" ~"Mean Jun-Aug Air Temperature, C", 
    metricName == "atDegDay" ~"# days with mean Air temperature >25 C"
  ))%>%
  mutate(site = str_remove(siteParam, "Air Temperature"), 
         siteMetric = paste(metricPrint, site, sep = "- "))

atPVs <- airTempPV%>%
  group_by(siteMetric, metricPrint)%>%
  summarize(x = mean(year), label = round(unique(p.value), 2))
```

Draw plots: 

```{r}
at.pal<-c("#E0F3DB","#CCEBC5", "#A8DDB5", "#7BCCC4","#4EB3D3","#2B8CBE")

at.plot <- ggplot(airTempPV, aes(x = year, y = metricValue, color = metricPrint))+
  geom_point()+
  geom_smooth(se = T, method = "lm", linetype = "dashed")+
  theme_classic()+
  labs(x = "Year", y = "Metric Value", color = "Metric")+
  scale_color_manual(values = at.pal)+
  geom_text(data = atPVs, aes(x = x, label = paste("P=", label)), y = Inf, vjust = 2)+
  facet_wrap(~siteMetric, scales = "free", labeller = label_wrap_gen(width = 40))+
  theme(axis.text = element_text(size = 11), 
        axis.title = element_text(size = 11, face = "bold"), 
        strip.text = element_text(size = 10, face = "bold"), 
        legend.position = "bottom")


at.plot

#Save plot:
ggsave("AirTemp_metrics.jpg", at.plot, device = "jpeg", path = here("outputs", "plots"), units = "in", dpi = "retina", height = 10, width = 10)
```

#### v. SWE:
  + Peak SWE
  + Date of peak SWE
  + "Meltout time": # days between peak SWE and zero SWE
  
First, extract data and calculate annual peak swe, date of peak swe, and # days between peak SWE and zero SWE; reorganize for analysis and plotting:

```{r}
snow <- loganSnotel%>%
  mutate(year = year(date))%>%
  group_by(site_name, year, description)%>%
  summarise(sweMax = max(snow_water_equivalent, na.rm = T), 
            sweMaxDOY = yday(date[which.max(snow_water_equivalent)]), 
            sweZero = yday(date[which.min(snow_water_equivalent)][1]), 
            nDays = length(unique(date)))%>%
  filter(!nDays<365)%>%
  mutate(meltDays = sweZero-sweMaxDOY)%>%
  mutate(meltDays = ifelse(meltDays<0, NA, meltDays))%>%
  select(-c(nDays,sweZero))%>%
  pivot_longer(!c(site_name, year, description), names_to = "metricName", values_to = "metricValue")
```

Nest data for modeling:  

```{r}
snowNest <- snow %>%
  nest(data = -c(site_name, metricName)) #create new nested dataframe with columns for site and metricName, one column for data - data entries for each site include year and metricValues. 
```

Run basic models for each site and metric of interest: 

```{r}
snowModels <- snowNest %>%
  mutate(model = map(data, ~lm(metricValue~year, data = .)%>% #run a model of temp~year for each site (referenced by "."); store results in new column called "model"
                     tidy))%>% #tidy function to construct a summary table with model results (seperate tables for each site)
  unnest(model) %>% #unnest model column to get rows for each site, columns for model metrics
  filter(term == "year")  #remove intercept term to just get info on year for each site
snowModels
```

Add P-values for plotting: 

```{r}
snowPV <- merge(snow, 
                     select(snowModels,site_name, metricName, p.value)
                     )%>%
  mutate(metricPrint = case_when(
    metricName == "sweMax"~ "Maximum Snow Water Equivalent, cm", 
    metricName == "sweMaxDOY" ~"Max Snow Water Equivalent Day of Year, C", 
    metricName == "meltDays" ~"# days between peak SWE and 0 SWE"
  ))%>%
  mutate(siteMetric = paste(site_name, metricPrint, sep = "- "))

sPVs <- snowPV%>%
  group_by(siteMetric, metricPrint)%>%
  summarize(x = mean(year), label = round(unique(p.value), 2))
```

Draw plots: 

```{r}
snow.plot <- ggplot(snowPV, aes(x = year, y = metricValue, color = metricPrint))+
  geom_point()+
  geom_smooth(se = T, method = "lm", linetype = "dashed")+
  theme_classic()+
  labs(x = "Year", y = "Metric Value", color = "Metric")+
  scale_color_manual(values = c("cyan4", "darkblue", "darkgreen"))+
  geom_text(data = sPVs, aes(x = x, label = paste("P=", label)), y = Inf, vjust = 2)+
  facet_wrap(~siteMetric, scales = "free", labeller = label_wrap_gen(width = 40))+
  theme(axis.text = element_text(size = 11), 
        axis.title = element_text(size = 11, face = "bold"), 
        strip.text = element_text(size = 10, face = "bold"), 
        legend.position = "bottom")


snow.plot

#Save plot:
ggsave("Snow_metrics.jpg", snow.plot, device = "jpeg", path = here("outputs", "plots"), units = "in", dpi = "retina", height = 10, width = 10)
```

## V. SNOTEL vs. Streamflow Metrics

### A. How does Max SWE at each SNOTEL site correlate with total annual discharge? 

Data prep: 

```{r}
snotelMerge <- snow%>%
  select(site_name, year, metricName, metricValue)%>%
  filter(metricName == "sweMax")%>%
  pivot_wider(id_cols = year, names_from = site_name, values_from = metricValue)%>%
  rename(klondikeSWE = 2, 
         templeForkSWE = 3, 
         tonyGroveLakeSWE = 4, 
         tonyGroveRsSWE = 5) #data wrangling to get dataframe with row for each year, columns for max SWE reading at each snotel station

qMerge <- discharge%>%
  select(siteParam, year, metricName, metricValue)%>%
  ungroup()%>%
  filter(metricName == "annualTot")%>%
  mutate(site = str_remove(siteParam, "Discharge"))%>%
  select(-siteParam)%>%
  pivot_wider(id_cols = year, names_from = site, values_from = metricValue)%>%
  rename(
    loganAtMainCFS = 2, 
    loganAtMendonCFS = 3, 
    loganAtFbCFS = 4
  ) #data wrangling to get dataframe with row for each year, columns for total annual discharge readings for each LRO site

lroSnotel <- merge(snotelMerge, qMerge)%>%
  pivot_longer(-c(year, loganAtMainCFS, loganAtMendonCFS, loganAtFbCFS), names_to = "snotelSite", values_to = "maxSWE") #merge, pivot longer for analysis and plotting
```

**Analysis/basic LM's: How does each SNOTEL site correlate with discharge?**

Nest data: 

```{r}
stQNest <- lroSnotel %>%
  nest(data = -c(snotelSite)) #create new nested dataframe with columns for site and metricName, one column for data - data entries for each site include year and metricValues. 
```

#### i. SNOTEL sites vs. Logan River at Franklin Basin

Run model for each SNOTEL site vs. discharge at Franklin Basin:

```{r}
stQFb <- stQNest %>%
  mutate(model = map(data, ~lm(loganAtFbCFS~maxSWE, data = .)%>% #run a model of temp~year for each site (referenced by "."); store results in new column called "model"
                     glance))%>% #tidy function to construct a summary table with model results (seperate tables for each site)
  unnest(model) #%>% #unnest model column to get rows for each site, columns for model metrics
  #filter(term == "maxSWE")  #remove intercept term to just get info on year for each site
stQFb
```
Add P-values for plotting: 

```{r}
stQFbPV <- merge(lroSnotel, 
                     select(stQFb, snotelSite, adj.r.squared, p.value)
                     )

fbPVs <- stQFbPV%>%
  group_by(snotelSite)%>%
  summarize(x = mean(loganAtFbCFS), p.label = round(unique(p.value), 2), r.label = round(unique(adj.r.squared),2))
```

Draw plots: 

```{r}
fb.plot <- ggplot(stQFbPV, aes(x = loganAtFbCFS, y = maxSWE, color = snotelSite))+
  geom_point()+
  geom_smooth(se = T, method = "lm", linetype = "dashed")+
  theme_classic()+
  labs(x = "Logan River at Franklin Basin - Total Annual Discharge, CFS", y = "Peak Snow Water Equivalent, cm", color = "Snotel Site")+
  scale_color_manual(values = brewer.pal(length(unique(stQFbPV$snotelSite)), "Set3"))+
  geom_text(data = fbPVs, aes(x = x, label = paste("P=", p.label, ", ", "R2=", r.label)), y = Inf, vjust = 2)+
  facet_wrap(~snotelSite, scales = "free", labeller = label_wrap_gen(width = 40))+
  theme(axis.text = element_text(size = 11), 
        axis.title = element_text(size = 11, face = "bold"), 
        strip.text = element_text(size = 10, face = "bold"), 
        legend.position = "none")

fb.plot


#Save plot:
ggsave("FranklinBasin_SNOTEL.jpg", fb.plot, device = "jpeg", path = here("outputs", "plots"), units = "in", dpi = "retina", height = 10, width = 10)
```

#### ii. SNOTEL sites vs. Logan River at Main St

Run model for each SNOTEL site vs. discharge at Main St.:

```{r}
stQMain <- stQNest %>%
  mutate(model = map(data, ~lm(loganAtMainCFS~maxSWE, data = .)%>% #run a model of temp~year for each site (referenced by "."); store results in new column called "model"
                     glance))%>% #tidy function to construct a summary table with model results (seperate tables for each site)
  unnest(model) #%>% #unnest model column to get rows for each site, columns for model metrics
  #filter(term == "maxSWE")  #remove intercept term to just get info on year for each site
stQMain
```
Add P-values for plotting: 

```{r}
stQMainPV <- merge(lroSnotel, 
                     select(stQMain, snotelSite, p.value, adj.r.squared)
                     )

mainPVs <- stQMainPV%>%
  group_by(snotelSite)%>%
  summarize(x = mean(loganAtMainCFS), p.label = round(unique(p.value), 2), r.label=round(unique(adj.r.squared), 2))
```

Draw plots: 

```{r}
main.plot <- ggplot(stQMainPV, aes(x = loganAtMainCFS, y = maxSWE, color = snotelSite))+
  geom_point()+
  geom_smooth(se = T, method = "lm", linetype = "dashed")+
  theme_classic()+
  labs(x = "Logan River at Main St. - Total Annual Discharge, CFS", y = "Peak Snow Water Equivalent, cm", color = "Snotel Site")+
  scale_color_manual(values = brewer.pal(length(unique(stQFbPV$snotelSite)), "Set3"))+
  geom_text(data = mainPVs, aes(x = x, label = paste("P=", p.label, ", ", "R2=", r.label)), y = Inf, vjust = 2)+
  facet_wrap(~snotelSite, scales = "free", labeller = label_wrap_gen(width = 40))+
  theme(axis.text = element_text(size = 11), 
        axis.title = element_text(size = 11, face = "bold"), 
        strip.text = element_text(size = 10, face = "bold"), 
        legend.position = "none")


main.plot

#Save plot:
ggsave("MainSt_SNOTEL.jpg", main.plot, device = "jpeg", path = here("outputs", "plots"), units = "in", dpi = "retina", height = 10, width = 10)
```

#### iii. SNOTEL sites vs. Logan River at Mendon

Run model for each SNOTEL site vs. discharge at Mendon:

```{r}
stQMendon <- stQNest %>%
  mutate(model = map(data, ~lm(loganAtMendonCFS~maxSWE, data = .)%>% #run a model of temp~year for each site (referenced by "."); store results in new column called "model"
                     glance))%>% #tidy function to construct a summary table with model results (seperate tables for each site)
  unnest(model) #%>% #unnest model column to get rows for each site, columns for model metrics
  #filter(term == "maxSWE")  #remove intercept term to just get info on year for each site
stQMendon
```

Add P-values for plotting: 

```{r}
stQMendonPV <- merge(lroSnotel, 
                     select(stQMendon, snotelSite, p.value, adj.r.squared)
                     )

mendonPVs <- stQMendonPV%>%
  group_by(snotelSite)%>%
  summarize(x = mean(loganAtMendonCFS), p.label = round(unique(p.value), 2), r.label = round(unique(adj.r.squared), 2))
```

Draw plots: 

```{r}
mendon.plot <- ggplot(stQFbPV, aes(x = loganAtMendonCFS, y = maxSWE, color = snotelSite))+
  geom_point()+
  geom_smooth(se = T, method = "lm", linetype = "dashed")+
  theme_classic()+
  labs(x = "Logan River at Mendon - Total Annual Discharge, CFS", y = "Peak Snow Water Equivalent, cm", color = "Snotel Site")+
  scale_color_manual(values = brewer.pal(length(unique(stQFbPV$snotelSite)), "Set3"))+
  geom_text(data = mendonPVs, aes(x = x, label = paste("P=", p.label, ", ", "R2=", r.label)), y = Inf, vjust = 2)+
  facet_wrap(~snotelSite, scales = "free", labeller = label_wrap_gen(width = 40))+
  theme(axis.text = element_text(size = 11), 
        axis.title = element_text(size = 11, face = "bold"), 
        strip.text = element_text(size = 10, face = "bold"), 
        legend.position = "none")


mendon.plot

#Save plot:
ggsave("Mendon_SNOTEL.jpg", mendon.plot, device = "jpeg", path = here("outputs", "plots"), units = "in", dpi = "retina", height = 10, width = 10)
```

### B. How does max SWE at all sites together correlate with streamflow? 

Basic concept = multi-variate linear model with all four snotel sites for each discharge monitoring station. 

Data prep: 

```{r}
lroStWide <- lroSnotel%>%
  pivot_wider(id_cols = c(year, loganAtFbCFS, loganAtMainCFS, loganAtMendonCFS), names_from = snotelSite, values_from = maxSWE)
```

#### i. Multivariate model for Franklin Basin

Fit model: 

```{r}
fb.multivariate <- lm(loganAtFbCFS~klondikeSWE+templeForkSWE+tonyGroveLakeSWE+tonyGroveRsSWE, data = lroStWide)
```

Summary: 

```{r}
summary(fb.multivariate)
```

**Result:** multivariate model does much worse than individual sites (Klondike, Tony Grove RS) - no siginificant correlation. 

#### i. Multivariate model for Main Street

Fit model: 

```{r}
main.multivariate <- lm(loganAtMainCFS~klondikeSWE+templeForkSWE+tonyGroveLakeSWE+tonyGroveRsSWE, data = lroStWide)
```

Summary: 

```{r}
summary(main.multivariate)
```

**Result:** multivariate model worse than individual models. 

#### i. Multivariate model for Mendon

Fit model: 

```{r}
mendon.multivariate <- lm(loganAtMendonCFS~klondikeSWE+templeForkSWE+tonyGroveLakeSWE+tonyGroveRsSWE, data = lroStWide)
```

Summary: 

```{r}
summary(mendon.multivariate)
```

**Result:** Multivariate model does better than the other two, but still not as good as single sites on their own - only Tony Grove Lake is a significant predictor. 

### C. How does basin-wide (all SNOTEL sites) mean max SWE correlate with total annual Q at each site? 

Data prep: 

```{r}
lroStNest <- lroStWide%>%
  pivot_longer(-c(year, klondikeSWE, templeForkSWE, tonyGroveLakeSWE, tonyGroveRsSWE), names_to = "lroSite", values_to = "totalAnnualQ")%>% #pivot longer for easier analysis and plotting by LRO site
  group_by(year)%>%
  mutate(maxSWEXbar = mean(c(klondikeSWE, templeForkSWE, tonyGroveLakeSWE, tonyGroveRsSWE)))%>% #calculate mean Max SWE across all four SNOTEL sites
  ungroup()%>%
  nest(data=!lroSite) #ungroup, nest by LRO site for modeling
```

Modeling: 

```{r}
meanSWEmods <- lroStNest %>%
  mutate(model = map(data, ~lm(totalAnnualQ~maxSWEXbar, data = .)%>% #run a model of temp~year for each site (referenced by "."); store results in new column called "model"
                     tidy))%>% #tidy function to construct a summary table with model results (seperate tables for each site)
  unnest(model) %>% #unnest model column to get rows for each site, columns for model metrics
  filter(term == "maxSWEXbar")  #remove intercept term to just get info on year for each site
meanSWEmods
```

Add P-values, prep for plotting: 

```{r}
lroStPlot <- lroStNest%>%
  unnest()%>%
  merge(select(meanSWEmods, lroSite, p.value))

lroStPV <- lroStPlot%>%
  group_by(lroSite)%>%
  summarize(x = mean(totalAnnualQ), label = round(unique(p.value), 2))
```

Draw plots: 

```{r}
meanSwe.plot <- ggplot(lroStPlot, aes(x = totalAnnualQ, y = maxSWEXbar, color = lroSite))+
  geom_point()+
  geom_smooth(se = T, method = "lm", linetype = "dashed")+
  theme_classic()+
  labs(x = "Total Annual Discharge, CFS", y = "Mean Peak Snow Water Equivalent, cm", color = "Snotel Site")+
  scale_color_manual(values = brewer.pal(length(unique(lroStPlot$lroSite)), "PRGn"))+
  geom_text(data = lroStPV, aes(x = x, label = paste("P=", label)), y = Inf, vjust = 2)+
  facet_wrap(~lroSite, scales = "free", labeller = label_wrap_gen(width = 40))+
  theme(axis.text = element_text(size = 11), 
        axis.title = element_text(size = 11, face = "bold"), 
        strip.text = element_text(size = 10, face = "bold"), 
        legend.position = "none")


meanSwe.plot

#Save plot:
ggsave("MeanSWE_AnnualQ.jpg", meanSwe.plot, device = "jpeg", path = here("outputs", "plots"), units = "in", dpi = "retina", height = 10, width = 10)
```

## VI. SNOTEL sites relevancy analysis

Goal = check all Snotel sites in the greater logan area for effectiveness at predicting streamflow in the Logan. 

Data prep: 

```{r}
snotelAll_merge <- allSnotel%>%
  mutate(year = year(date))%>% #add year column
  group_by(site_name, year)%>%
  summarise(sweMax = max(snow_water_equivalent), 
            sweMean = mean(snow_water_equivalent)) #calculate max and mean SWE for each site/year
```

Merge with LRO discharge data:

```{r}
snotelAllLro <- merge(snotelAll_merge, qMerge)%>%
  mutate(siteCode = as.numeric(site_name))%>%
  arrange(site_name, year)%>%
  mutate(site_name = as.character(site_name))%>%
  filter(!site_name == "clarkston mountain ")

unique(snotelAllLro$site_name)
```


### A. All SNOTEL sites vs. Logan at Franklin Basin

Modeling with a for-loop: 

```{r}
sites <- unique(snotelAllLro$site_name)
FbModelResults <- tibble()

for(i in 1:length(sites)){
  d <- filter(snotelAllLro, site_name == sites[i])
  mod <- lm(loganAtFbCFS~sweMax, data = d)
  modSum <-tibble(sites[i], glance(mod)$adj.r.squared, glance(mod)$p.value)%>%
    rename(site_name = 1, adjR2 = 2, pVal = 3)

  FbModelResults<-rbind(modSum, FbModelResults)
  
  
}
```

Add R2 values for plotting: 

```{r}
FbStPlot <- snotelAllLro%>%
  merge(select(FbModelResults, site_name, adjR2))

FbStR2 <- FbStPlot%>%
  group_by(site_name)%>%
  summarize(x = mean(loganAtFbCFS, na.rm = T), label = round(unique(adjR2), 2))
```

Draw plots:


```{r}
FbSt.Plot <- ggplot(FbStPlot, aes(x = loganAtFbCFS, y = sweMax, color = site_name))+
  geom_point()+
  geom_smooth(se = T, method = "lm", linetype = "dashed")+
  theme_classic()+
  labs(x = "Logan River at Franklin Basin -- Total Annual Discharge, CFS", y = "Peak Snow Water Equivalent, cm", color = "Snotel Site")+
  #scale_color_manual(values = c("cyan2", "cyan3", "cyan4"))+
  geom_text(data = FbStR2, aes(x = x, label = paste("R2=", label)), y = Inf, vjust = 2)+
  facet_wrap(~site_name, scales = "free", labeller = label_wrap_gen(width = 40))+
  theme(axis.text = element_text(size = 11), 
        axis.title = element_text(size = 11, face = "bold"), 
        strip.text = element_text(size = 10, face = "bold"), 
        legend.position = "none")


FbSt.Plot

#Save plot:
ggsave("Fb_SnotelAnalysis.jpg", FbSt.Plot, device = "jpeg", path = here("outputs", "plots"), units = "in", dpi = "retina", height = 10, width = 10)
```


### B. All SNOTEL sites vs. Logan at Main St

Modeling with a for-loop: 

```{r}
sites <- unique(snotelAllLro$site_name)
MainModelResults <- tibble()

for(i in 1:length(sites)){
  d <- filter(snotelAllLro, site_name == sites[i])
  mod <- lm(loganAtMainCFS~sweMax, data = d)
  modSum <-tibble(sites[i], glance(mod)$adj.r.squared, glance(mod)$p.value)%>%
    rename(site_name = 1, adjR2 = 2, pVal = 3)
  
  MainModelResults<-rbind(modSum, MainModelResults)
  
}

```

Add R2 values for plotting: 

```{r}
MainStPlot <- snotelAllLro%>%
  merge(select(MainModelResults, site_name, adjR2))

MainStR2 <- MainStPlot%>%
  group_by(site_name, na.rm = T)%>%
  summarize(x = mean(loganAtMainCFS), label = round(unique(adjR2), 2))
```

Draw plots:


```{r}
MainSt.Plot <- ggplot(MainStPlot, aes(x = loganAtMainCFS, y = sweMax, color = site_name))+
  geom_point()+
  geom_smooth(se = T, method = "lm", linetype = "dashed")+
  theme_classic()+
  labs(x = "Logan River at Main St. -- Total Annual Discharge, CFS", y = "Peak Snow Water Equivalent, cm", color = "Snotel Site")+
  #scale_color_manual(values = c("cyan2", "cyan3", "cyan4"))+
  geom_text(data = MainStR2, aes(x = x, label = paste("R2=", label)), y = Inf, vjust = 2)+
  facet_wrap(~site_name, scales = "free", labeller = label_wrap_gen(width = 40))+
  theme(axis.text = element_text(size = 11), 
        axis.title = element_text(size = 11, face = "bold"), 
        strip.text = element_text(size = 10, face = "bold"), 
        legend.position = "none")


MainSt.Plot

#Save plot:
ggsave("Main_SnotelAnalysis.jpg", MainSt.Plot, device = "jpeg", path = here("outputs", "plots"), units = "in", dpi = "retina", height = 10, width = 10)
```


### C. All SNOTEL sites vs. Logan at Mendon St

Modeling with a for-loop: 

```{r}
sites <- unique(snotelAllLro$site_name)
MendonModelResults <- tibble()

for(i in 1:length(sites)){
  d <- filter(snotelAllLro, site_name == sites[i])
  mod <- lm(loganAtMendonCFS~sweMax, data = d)
  modSum <-tibble(sites[i], glance(mod)$adj.r.squared, glance(mod)$p.value)%>%
    rename(site_name = 1, adjR2 = 2, pVal = 3)
  
  MendonModelResults<-rbind(modSum, MendonModelResults)

}
```

Add R2 values for plotting: 

```{r}
MendonStPlot <- snotelAllLro%>%
  merge(select(MendonModelResults, site_name, adjR2))

MendonStR2 <- MendonStPlot%>%
  group_by(site_name)%>%
  summarize(x = mean(loganAtMendonCFS, na.rm = T), label = round(unique(adjR2), 2))
```

Draw plots:


```{r}
MendonSt.Plot <- ggplot(MendonStPlot, aes(x = loganAtMendonCFS, y = sweMax, color = site_name))+
  geom_point()+
  geom_smooth(se = T, method = "lm", linetype = "dashed")+
  theme_classic()+
  labs(x = "Logan River at Mendon -- Total Annual Discharge, CFS", y = "Peak Snow Water Equivalent, cm", color = "Snotel Site")+
  #scale_color_manual(values = c("cyan2", "cyan3", "cyan4"))+
  geom_text(data = MendonStR2, aes(x = x, label = paste("R2=", label)), y = Inf, vjust = 2)+
  facet_wrap(~site_name, scales = "free", labeller = label_wrap_gen(width = 40))+
  theme(axis.text = element_text(size = 11), 
        axis.title = element_text(size = 11, face = "bold"), 
        strip.text = element_text(size = 10, face = "bold"), 
        legend.position = "none")


MendonSt.Plot

#Save plot:
ggsave("Mendon_SnotelAnalysis.jpg", MendonSt.Plot, device = "jpeg", path = here("outputs", "plots"), units = "in", dpi = "retina", height = 10, width = 10)
```

## VII. Updated Analyses - Key Metrics for each site


### A. Snowpack

Key indicators of snowpack are: 

  - Peak SWE amount 
  - Peak SWE date
  - Meltout timing (# days between peak SWE and zero SWE)

Data needed for these variables: Need SNOTEL data for Jan 1 - June 1, should exclude years without data. 


#### i. Peak SWE Amount, Date, and Meltout timing

QA/QC: Extract SNOTEL data from Jan 1 to July 1 of each year; drop years with missing data during that time period

Starting with key SNOTEL sites: 

```{r}
# Drop partial years
SWEPeaks <- loganSnotel%>%
  mutate(month = month(date),
         year = year(date),
         doy = yday(date))%>% #add month and day of year columns
  mutate(waterYear = ifelse(month>=10, year+1, year))%>% #add water year column
  group_by(site_name,year)%>%
  mutate(doyMin = min(doy), 
         doyMax = max(doy))%>% #calculate min and max DOY for each site/year
  mutate(waterDOY = ifelse(doy>=doyMax-91, 
                           doy-(doyMax-92), 
                           doy+92)
         )%>% # #use min/max Doy to calculate water day of year
  group_by(site_name, waterYear)%>%
  mutate(waterDoyMax = max(waterDOY), 
         waterDoyMin = min(waterDOY))%>%
  #filter(!(doyMin>1|waterDoyMax<274))%>% #drop years that start after Jan 1st or end before July 1st
  select(site_name, date, year, doy,waterYear, waterDOY, month, snow_water_equivalent, waterDoyMin, waterDoyMax, doyMin, doyMax)#drop extra cols
  

#Check for missing data:
print(paste("Number of NAs in dataframe:", length(which(is.na(SWEPeaks$snow_water_equivalent))), sep = " "))

#Appear to be very few missing values. Drop from dataframe: 

SWEPeaks <- SWEPeaks %>%
  filter(!is.na(snow_water_equivalent)) #drop missing data
```

Next, calculate peak SWE amount, peak SWE date, and meltout days for each water year at each site

```{r}
sweStats <- SWEPeaks%>%
  filter(!doyMin>1, !doyMax<182)%>% #drop years that start after jan 1 or end before jul 1
  filter(doy>1&doy<182)%>% #extract Jan 1 to July 1
  group_by(site_name, waterYear)%>% #group by site and year
  summarise(sweMax = max(snow_water_equivalent, na.rm = T), 
            sweMaxDOY = yday(date[which.max(snow_water_equivalent)]), 
            sweZero = yday(date[which.min(snow_water_equivalent)][1]))%>% #calculate maximum SWE value, date of max SWE, and first day with 0 SWE recorded
  mutate(meltDays = sweZero-sweMaxDOY)%>% #Calculate melt day duration by subtracting max SWE DOY from DOY of first day with 0 SWE
 pivot_longer(!c(site_name, waterYear), names_to = "metricName", values_to = "metricValue")
```

Basic LM for each metric over time: 

```{r}
sweStatsNest <- sweStats %>%
  nest(data = -c(site_name, metricName)) #create new nested dataframe with columns for site and metricName, one column for data - data entries for each site include year and metricValues. 
```

Run basic models for each site and metric of interest: 

```{r}
sModels <- sweStatsNest %>%
  mutate(model = map(data, ~lm(metricValue~waterYear, data = .)%>% #run a model of temp~year for each site (referenced by "."); store results in new column called "model"
                     tidy))%>% #tidy function to construct a summary table with model results (seperate tables for each site)
  unnest(model) %>% #unnest model column to get rows for each site, columns for model metrics
  filter(term == "waterYear")  #remove intercept term to just get info on year for each site
sModels
```

Add P-values for plotting: 

```{r}
swePV <- merge(sweStats, 
                     select(sModels, site_name, metricName, p.value)
                     )%>%
  mutate(metricPrint = case_when(
    metricName == "sweMax" ~ "Maximum Recorded SWE, mm", 
    metricName == "sweMaxDOY" ~"Day of year of max. recorded SWE", 
    metricName == "sweZero" ~"First day of year w/ zero recorded SWE",
    metricName == "meltDays" ~"Melt duration, days"
  ))%>%
  mutate(siteMetric = paste(site_name, metricPrint, sep = " - "))

sPVs <- swePV%>%
  group_by(siteMetric, metricPrint)%>%
  summarize(x = mean(waterYear), label = round(unique(p.value), 2))
```

Plotting:

```{r}
s.plot <- ggplot(swePV, aes(x = waterYear, y = metricValue, color = metricPrint))+
  geom_point()+
  geom_smooth(se = T, method = "lm", linetype = "dashed")+
  theme_classic()+
  labs(x = "Water Year", y = "Metric Value", color = "Metric")+
  scale_color_manual(values = c("cyan4", "darkblue", "darkgreen", "cyan3"))+
  geom_text(data = sPVs, aes(x = x, label = paste("P=", label)), y = Inf, vjust = 2)+
  facet_wrap(~siteMetric, scales = "free", labeller = label_wrap_gen(width = 40))+
  theme(axis.text = element_text(size = 11), 
        axis.title = element_text(size = 11, face = "bold"), 
        strip.text = element_text(size = 10, face = "bold"), 
        legend.position = "bottom")


s.plot

#Save plot:
ggsave("SWE_metrics.jpg", s.plot, device = "jpeg", path = here("outputs", "plots", "updatedTrends"), units = "in", dpi = "retina", height = 10, width = 10)
```

### B. Precip

Key indicators for precip: 
  - Total water year precip (OR total annual precip)
  - Amount of precip, rain vs. snow

Data Needs: 

**For cumulative precip:** Need precip data for entire year - exclude partial years from analysis. Need to get rid of negative cumulative precip values first. 

**For rain vs. snow**, Need Precip Data + Air temp data. Use empirical formula from Dai et al. 2008 (referenced in SI from YSL paper). 

**Formulas:** 

Snow Probability = -50 * [tanh(0.4 * ( Mean Temp – 1.75))-1]

WHERE: Mean Temp = mean of the maximum and minimum tempreatures on each day

Snowfall amount = Snow Probability * Daily Precip total

Rainfall amount = (1 - Snow Probability) * Daily Precip total OR Daily Precip total - Snowfall amount

**Refrence**:

Dai, A. 2008. Temperature and pressure dependence of the rain-snow phase transition over
land and ocean. Geophys. Res. Lett. 35: L12802, doi:10.1029/2008GL033295

#### i. Cumulative Precip

*QA/QC:* Clean up data, merge LRO and SNOTEL datasets 

  - Looks like the "year" for cumulative calculations re-sets at 7am on 1/1 of each year, not midnight - need to deal with these incorrect values. 
  - Also have some negative cumulative precip values early in year
  - When calculating measured precip, get some negative values - indicates error with cumulative precip. Need to get rid of errors and re-calculate precip cumulative. 

```{r}
#LRO Precip QA/QC:
lroClimate <- fullData_complete%>%
  filter(str_detect(site, "Climate"))%>% #Extract climate stations
  select(dateTime, parameterValue, site, parameter)%>% #drop extra columns
  pivot_wider(id_cols = c(dateTime, site), names_from = parameter, values_from = parameterValue)%>% #pivot wider to get columns for each variable
  rename(site_name = 2, precipitation_cumulative = 4, airTemp = 3)%>% #renaming for merge
  mutate(year = year(dateTime), doy = yday(dateTime), time = format(as.POSIXct(dateTime), "%H:%M:%S"))%>% #adding time, year, and doy columns
  mutate(precipitation_cumulative = ifelse(doy==1&time<"07:00:00", 0, precipitation_cumulative))%>% #replacing carryover totals from prev. year with 0
 mutate(precipitation_cumulative = ifelse(precipitation_cumulative<0, 0, precipitation_cumulative))%>% #replace negative cumulative precip values with 0
  filter(!is.na(precipitation_cumulative))%>% #drop rows without cumulative precip values
  group_by(year, site_name)%>% # group by year and site
  mutate(precipitation = (precipitation_cumulative-lag(precipitation_cumulative)))%>% #calculate measured precip (current cumulative value minus previous cumulative value)
  ungroup()%>%
  mutate(precipitation = ifelse((is.na(precipitation)|precipitation<0), 0, precipitation))%>% #replace negative measured precip values with 0
  group_by(year, site_name)%>% #group by year and site name again
  mutate(precipCumulative_cor = cumsum(precipitation))%>% #re-calculate cumulative precip for each year with corrected daily precip values
  mutate(precipCumulative_cor = ifelse(is.na(precipCumulative_cor), 0, precipCumulative_cor)) #Replace NA's created from cumsum fxn with 0

#checking output
check <- lroClimate%>%
  group_by(site_name, year)%>% #group by site name and year
  summarise(stated = max(precipitation_cumulative, na.rm = T), 
            calc = sum(precipitation), 
            cor = max(precipCumulative_cor))%>% #calculate stated, calculated, and corrected cumulative precip
  mutate(diff = calc-stated, 
         diffCor = calc-cor) #calculate differences between calculated and stated/corrected values

#re-organize + calculations for merge with Snotel data: 

lroClimateMerge <- lroClimate%>%
  select(dateTime, site_name, airTemp, precipitation, precipCumulative_cor)%>%
  mutate(date = ymd(substr(dateTime, 1, 10)))%>%
  group_by(date, site_name)%>%
  summarize(
    precipitation_cumulative = (max(precipCumulative_cor)*10), 
    precipitation = (sum(precipitation)*10), 
    temperature_max = max(airTemp), 
    temperature_min = min(airTemp), 
    temperature_mean = mean(airTemp)
  )%>%
  arrange(site_name, date)

#Extract climate data columns from SNOTEL data:

snotelClimate <- loganSnotel%>%
  select(site_name, date, precipitation_cumulative, precipitation, temperature_max, temperature_min, temperature_mean)

#Merge: 

precipFull <- rbind(lroClimateMerge, snotelClimate)%>%
  mutate(temperature_mean = ifelse(temperature_mean == -99.9, NA, temperature_mean))
```

*Cumulative Precip Analysis:*

Extract cumulative precip for each year; nest by site

```{r}
precipStats <- precipFull %>%
  mutate(year = year(date))%>%
  group_by(site_name, year)%>%
  summarize(totPrecip = max(precipitation_cumulative))

precipNest<-precipStats%>%
  nest(data = -c(site_name))
```

Run basic models for each site and metric of interest: 

```{r}
pModels <- precipNest %>%
  mutate(model = map(data, ~lm(totPrecip~year, data = .)%>% #run a model of temp~year for each site (referenced by "."); store results in new column called "model"
                     tidy))%>% #tidy function to construct a summary table with model results (seperate tables for each site)
  unnest(model) %>% #unnest model column to get rows for each site, columns for model metrics
  filter(term == "year")  #remove intercept term to just get info on year for each site
pModels
```

Add P-values for plotting: 

```{r}
precPV <- merge(precipStats, 
                     select(pModels, site_name, p.value)
                     )

pPVs <- precPV%>%
  group_by(site_name)%>%
  summarize(x = mean(year), label = round(unique(p.value), 2))
```

Plotting:

```{r}
p.plot <- ggplot(precPV, aes(x = year, y = totPrecip, color = site_name))+
  geom_point()+
  geom_smooth(se = T, method = "lm", linetype = "dashed")+
  theme_classic()+
  labs(x = "Year", y = "Cumulative annual precip, mm", color = "Site")+
  scale_color_manual(values = c("#7FCDBB" ,"#41B6C4" ,"#1D91C0" ,"#225EA8" ,"#253494" ,"#081D58"))+
  geom_text(data = pPVs, aes(x = x, label = paste("P=", label)), y = Inf, vjust = 2)+
  facet_wrap(~site_name, scales = "free", labeller = label_wrap_gen(width = 40))+
  theme(axis.text = element_text(size = 11), 
        axis.title = element_text(size = 11, face = "bold"), 
        strip.text = element_text(size = 10, face = "bold"), 
        legend.position = "none")


p.plot

#Save plot:
ggsave("cumulative_precip.jpg", p.plot, device = "jpeg", path = here("outputs", "plots", "updatedTrends"), units = "in", dpi = "retina", height = 10, width = 10)
```

#### ii. Snow vs. Rain

Data prep & QA/QC - extract data with both air temp and precip; use empirical formula (Snow Probability = -50 * [tanh(0.4 * ( Mean Temp – 1.75))-1]) to calculate snow ratio: 

```{r}
rainSnow <- precipFull%>%
  filter(!(is.na(temperature_max)|is.na(temperature_min)))%>% #drop rows without max and min temp
  mutate(temperature_mean = ((temperature_max+temperature_min)/2))%>% #re-calc mean daily temp to ensure consistency
  mutate(snowProb = round(-50*(tanh(0.4*(temperature_mean-1.75))-1), 3)/100
         )%>% #calculate snow probability using  method from Dai 2008, parameters from Tronstead et al. 2024. NOTE - had to divide by 100 to get numbers that made any sense. 
  mutate(dailySnow = precipitation*snowProb, #calculate daily snow amount as daily precip * snow probability
         dailyRain = precipitation*(1-snowProb))%>% #calculate daily rain amount as daily precip * (1-snow probability). Assumes no sleet. 
  mutate(precipCheck = dailySnow+dailyRain) #add column to check that rain amount + snow amount = daily precip amount. 
```

Calculate yearly summary for each site/year:

```{r}
rainSnowSum <- rainSnow%>%
  mutate(year = year(date),
         month = month(date))%>% #extract date and year 
  mutate(waterYear = ifelse(month<10, year, year+1))%>% #calculate water year
  group_by(site_name, waterYear)%>%
  summarise(
    totSnow = sum(dailySnow, na.rm = T), 
    totRain = sum(dailyRain, na.rm = T), 
    rainSnowRatio = (sum(dailyRain, na.rm = T)/sum(dailySnow, na.rm = T)) #calculate total snow, total rain, and rain:snow ratio for each water year/site
  )%>%
  pivot_longer(!c(site_name, waterYear), names_to = "metricName", values_to = "metricValue") #pivot longer for analysis and plotting. 
```

Basic LM for each metric over time: 

```{r}
rainSnowNest <- rainSnowSum %>%
  nest(data = -c(site_name, metricName)) #create new nested dataframe with columns for site and metricName, one column for data - data entries for each site include year and metricValues. 
```

Run basic models for each site and metric of interest: 

```{r}
rsModels <- rainSnowNest %>%
  mutate(model = map(data, ~lm(metricValue~waterYear, data = .)%>% #run a model of temp~year for each site (referenced by "."); store results in new column called "model"
                     tidy))%>% #tidy function to construct a summary table with model results (seperate tables for each site)
  unnest(model) %>% #unnest model column to get rows for each site, columns for model metrics
  filter(term == "waterYear")  #remove intercept term to just get info on year for each site
rsModels
```

Add P-values for plotting: 

```{r}
rainSnowPV <- merge(rainSnowSum, 
                     select(rsModels, site_name, metricName, p.value) #add p-values
                     )%>%
  mutate(metricPrint = case_when(
    metricName == "totSnow" ~ "Total water year snowfall, mm", 
    metricName == "totRain" ~"Total water year rainfall, mm", 
    metricName == "rainSnowRatio" ~"Water year rain:snow ratio" #add print-friendly metric name for plotting
  ))%>%
  mutate(siteMetric = paste(site_name, metricPrint, sep = " - ")) #add metric + site column for plotting

rsPVs <- rainSnowPV%>%
  group_by(siteMetric, metricPrint)%>%
  summarize(x = mean(waterYear), label = round(unique(p.value), 2)) #create separate dataframe with p-values for adding to plots
```

Plotting:

```{r}
rs.plot <- ggplot(rainSnowPV, aes(x = waterYear, y = metricValue, color = metricPrint))+
  geom_point()+
  geom_smooth(se = T, method = "lm", linetype = "dashed")+
  theme_classic()+
  labs(x = "Water Year", y = "Metric Value", color = "Metric")+
  scale_color_manual(values = c("darkgreen","cyan4", "darkblue"))+
  geom_text(data = rsPVs, aes(x = x, label = paste("P=", label)), y = Inf, vjust = 2)+
  facet_wrap(~siteMetric, scales = "free", labeller = label_wrap_gen(width = 40))+
  theme(axis.text = element_text(size = 11), 
        axis.title = element_text(size = 11, face = "bold"), 
        strip.text = element_text(size = 10, face = "bold"), 
        legend.position = "bottom")


rs.plot

#Save plot:
ggsave("RainSnow_metrics.jpg", rs.plot, device = "jpeg", path = here("outputs", "plots", "updatedTrends"), units = "in", dpi = "retina", height = 10, width = 15)
```

##### a. Winter snow vs. rain

Same analysis as above, but extracting data for November-May

```{r}
rainSnow <- precipFull%>%
  filter(!(is.na(temperature_max)|is.na(temperature_min)))%>% #drop rows without max and min temp
  mutate(temperature_mean = ((temperature_max+temperature_min)/2))%>% #re-calc mean daily temp to ensure consistency
  mutate(snowProb = round(-50*(tanh(0.4*(temperature_mean-1.75))-1), 3)/100
         )%>% #calculate snow probability using  method from Dai 2008, parameters from Tronstead et al. 2024. NOTE - had to divide by 100 to get numbers that made any sense. 
  mutate(dailySnow = precipitation*snowProb, #calculate daily snow amount as daily precip * snow probability
         dailyRain = precipitation*(1-snowProb))%>% #calculate daily rain amount as daily precip * (1-snow probability). Assumes no sleet. 
  mutate(precipCheck = dailySnow+dailyRain) #add column to check that rain amount + snow amount = daily precip amount. 
```

Calculate yearly summary for each site/year:

```{r}
rainSnowWinter <- rainSnow%>%
  mutate(year = year(date),
         month = month(date))%>% #extract date and year 
  mutate(waterYear = ifelse(month<10, year, year+1))%>% #calculate water year
  filter(month>11|month<4)%>% #extract data for November-May
  group_by(site_name, waterYear)%>%
  summarise(
    totSnow = sum(dailySnow, na.rm = T), 
    totRain = sum(dailyRain, na.rm = T), 
    rainSnowRatio = (sum(dailyRain, na.rm = T)/sum(dailySnow, na.rm = T)) #calculate total snow, total rain, and rain:snow ratio for each water year/site
  )%>%
  pivot_longer(!c(site_name, waterYear), names_to = "metricName", values_to = "metricValue") #pivot longer for analysis and plotting. 
```

Basic LM for each metric over time: 

```{r}
rainSnowWNest <- rainSnowWinter %>%
  nest(data = -c(site_name, metricName)) #create new nested dataframe with columns for site and metricName, one column for data - data entries for each site include year and metricValues. 
```

Run basic models for each site and metric of interest: 

```{r}
rswModels <- rainSnowWNest %>%
  mutate(model = map(data, ~lm(metricValue~waterYear, data = .)%>% #run a model of temp~year for each site (referenced by "."); store results in new column called "model"
                     tidy))%>% #tidy function to construct a summary table with model results (seperate tables for each site)
  unnest(model) %>% #unnest model column to get rows for each site, columns for model metrics
  filter(term == "waterYear")  #remove intercept term to just get info on year for each site
rswModels
```

Add P-values for plotting: 

```{r}
rainSnowWPV <- merge(rainSnowWinter, 
                     select(rswModels, site_name, metricName, p.value) #add p-values
                     )%>%
  mutate(metricPrint = case_when(
    metricName == "totSnow" ~ "Total Dec-Mar snowfall, mm", 
    metricName == "totRain" ~"Total Dec-Mar rainfall, mm", 
    metricName == "rainSnowRatio" ~"Dec-Mar rain:snow ratio" #add print-friendly metric name for plotting
  ))%>%
  mutate(siteMetric = paste(site_name, metricPrint, sep = " - ")) #add metric + site column for plotting

rswPVs <- rainSnowWPV%>%
  group_by(siteMetric, metricPrint)%>%
  summarize(x = mean(waterYear), label = round(unique(p.value), 2)) #create separate dataframe with p-values for adding to plots

unique(rainSnowWPV$metricPrint)
```

Plotting:

```{r}
rsw.plot <- ggplot(rainSnowWPV, aes(x = waterYear, y = metricValue, color = metricPrint))+
  geom_point()+
  geom_smooth(se = T, method = "lm", linetype = "dashed")+
  theme_classic()+
  labs(x = "Water Year", y = "Metric Value", color = "Metric")+
  scale_color_manual(values = c("darkgreen","cyan4", "darkblue"))+
  geom_text(data = rswPVs, aes(x = x, label = paste("P=", label)), y = Inf, vjust = 2)+
  facet_wrap(~siteMetric, scales = "free", labeller = label_wrap_gen(width = 40))+
  theme(axis.text = element_text(size = 11), 
        axis.title = element_text(size = 11, face = "bold"), 
        strip.text = element_text(size = 10, face = "bold"), 
        legend.position = "bottom")


rsw.plot

#Save plot:
ggsave("RainSnow_WinterMetrics.jpg", rsw.plot, device = "jpeg", path = here("outputs", "plots", "updatedTrends"), units = "in", dpi = "retina", height = 10, width = 15)
```

### C. Canyon Sites - Discharge Metrics

Canyon sites are: Logan River at Franklin Basin, Wood Camp Spring, Rick's Spring, Logan River at Tony Grove RS, Logan River at Guinavah,  Logan River at First Dam, right hand fork, and temple fork. 

Key metrics for canyon sites are: 

  - Runoff start date
  - Date of peak runoff
  - Peak runoff value
  - Runoff volume from Runoff start date thru Oct 1st. 
  
#### i. Data prep & QA/QC:

Extract canyon discharge data, prep for initial checks/plots:

```{r}
canyonSites <- fullData_usgs%>%
  filter(site == "Logan River near Franklin Basin"|site == "Wood Camp Spring"|site == "Logan River at First Dam"|site == "Ricks Spring above confluence with Logan River"|site == "Logan River at Guinavah Campground Bridge"|site == "Logan River near Tony Grove"|site == "Temple Fork above confluence with Logan River"|site == "Right Hand Fork above confluence with Logan River")%>% #extract canyon sites
  filter(parameter == "Discharge_cfs")%>% #extract discharge data
  select(!c(parameter, unit))%>% #drop extra cols
  rename(discharge_cfs = parameterValue)%>% #renaming cols
  mutate(discharge_cfs = ifelse(discharge_cfs == 0, NA, discharge_cfs))%>%
  mutate(year = year(dateTime), 
         month = month(dateTime), 
         doy = yday(dateTime), 
         date = ymd(substr(dateTime, 1, 10)))%>% #extract year, month, and date
  group_by(site, date, year, month, doy)%>%
  summarise(discharge_cfs = mean(discharge_cfs, na.rm = T))%>% #group by date, calculate mean daily discharge
  mutate(waterYear=ifelse(month<10, year, year+1), 
         logDischarge = log10(discharge_cfs))%>%  #add water year column, add log10 discharge column.
  group_by(site,year)%>% 
  mutate(doyMin = min(doy), 
         doyMax = max(doy))%>% #calculate min and max DOY for each site/year
  mutate(waterDOY = ifelse(doy>=doyMax-91, 
                           doy-(doyMax-92), 
                           doy+92)
         ) #use min/max Doy to calculate water day of year 


summary(canyonSites$discharge_cfs)
```

Initial plot to check data availability: 

```{r}
ggplot(canyonSites, aes(x = date, y = discharge_cfs))+
  geom_line()+
  theme_classic()+
  #scale_y_continuous(trans = "log10")+
  facet_wrap(~site, scales = "free")
```

Spotty data for springs, looks like missing parts of peaks in some years. Quick summary of data availability:

```{r}
dataAv <- canyonSites %>% 
  filter(!is.na(discharge_cfs))%>%
  group_by(site, year)%>%
  summarise(
    dateMin = min(date, na.rm = T), 
    dateMax = max(date, na.rm = T), 
    doyMin = min(doy), 
    doyMax = max(doy)
  )
dataAv
```

#### ii. Identify start of runoff for each site/year:

Trying to find changepoints - test with Franklin Basin: 


```{r}
fbTest <- canyonSites%>%
  filter(site=="Logan River near Franklin Basin"&date>="2022-01-01"&date<="2022-10-01")%>% #filter out data for Franklin Basin between Jan 1 and Oct 1 2023
  filter(!is.na(discharge_cfs))

#Plot to check for obvious step change(s):

ggplot(fbTest, aes(x=doy, y =discharge_cfs))+
  geom_line()+
  theme_classic()
```

Attempt using cpop function:

```{r}
fb23.cpop <- cpop::cpop(fbTest$discharge_cfs, fbTest$doy, #providing y and x vectors for cpop function
                    beta =80) #adjusting beta value to increase penalty for adding more changepoints (default of 2*log(length(y)) (~9.8) was returning CP's before major step-change)
```

Check results - plot, extract changepoint:

```{r}
plot(fb23.cpop)#+
  #labs(title = unique(fbTest$site))
changepoints(fb23.cpop)
changepoints(fb23.cpop)[1,1]
```

First CP with beta set to 80 seems to be working well for years with good data availability, but will need to get rid of years with incomplete data. For now, dropping years that start after DOY 60 (March 1st)

```{r}
canyonSites_complete <- canyonSites%>%
  mutate(siteYr = paste(site, year, sep = " "))%>% #add site year column for loop. 
  mutate(discharge_cfs = ifelse(is.nan(discharge_cfs), NA, discharge_cfs))%>%
  filter(!is.na(discharge_cfs))%>% #drop rows with missing discharge values
  group_by(site, year)%>% #group by site and year
  mutate(doyMin = min(doy), 
         doyMax = max(doy))%>% #re-calculate DOY min and max with NA's removed
  ungroup()%>%
  filter(!doyMin>60, !doyMax<275) #drop years that start after march 1st or end before oct 1st. 
```

##### a. Iterate changepoint extraction over site/year with a for loop: 

For loop to extract CP's and save plots for each site/year: 

```{r}
site.yrs <- unique(canyonSites_complete$siteYr) #vector with unique values of siteYr for iteration
cps <- data.frame() #empty data frame to recieve changepoint data
cp.plots <- list() #empty list to recieve cp plots

#For loop: 

for(i in 1:length(site.yrs)){
  
  d<-canyonSites_complete%>%
    filter(siteYr == site.yrs[i])%>%#extract "ith" site/year combo
    filter(!is.na(discharge_cfs))
  
  c<-cpop(d$discharge_cfs, d$doy, beta = 800)#feeding temperature data and doy into cpop, set beta to 80
  cp<-changepoints(c)[1,1]#extract first changepoint
  cpp<-plot(c)+ #draw plot of cp estimations
    labs(title = site.yrs[i], x = "Day of year", y = "Discharge, cfs") #plot title + labels
  
  cpdf<-data.frame(unique(d$site), unique(d$year), cp) #create dataframe with site, year, and changepoint DOY
  cps<-rbind(cps, cpdf) #append to cps dataframe
  
  cp.plots[[i]] <- cpp #store plot as "ith" item in plot list
  
  print(paste("Completed ", i, "/", length(site.yrs), sep = ""))
}

cps <- cps%>%
  rename("site" = 1, "year" = 2, "cpDoy" = 3) #renaming cps columns for easier use later
```

Get references for plots for each site:

```{r}
siteRefs <- canyonSites_complete%>%
  select(site, siteYr)%>% #extract site and siteyr column
  distinct()%>% #drop duplicates
  group_by(site)%>%
  summarise(n = n()) #group by site, count # years
siteRefs
```

*Check FB plots:* 

```{r}
(cp.plots[[39]]|cp.plots[[40]]|cp.plots[[41]])/(cp.plots[[42]]|cp.plots[[43]]|cp.plots[[44]])/(cp.plots[[45]]|cp.plots[[46]]|cp.plots[[47]])
```

First CP looks good with beta at 800 for all years.  

*Check Ricks plots:* 

```{r}
(cp.plots[[48]]|cp.plots[[49]]|cp.plots[[50]])/(cp.plots[[51]]|cp.plots[[52]])
```

Beta at 800 works well. 

*Check Wood Camp plots:* 

```{r}
(cp.plots[[53]]/cp.plots[[54]])
```

Both look good with Beta at 800

*Spot Checking every 5yrs for First Dam*:

```{r}
(cp.plots[[1]]|cp.plots[[5]]|cp.plots[[10]])/(cp.plots[[15]]|cp.plots[[20]]|cp.plots[[25]])/(cp.plots[[30]]|cp.plots[[35]])
```

Beta at 800 works well, picks up actual peak instead of pre-peak squiggles. 

##### b. Add 1st change point data back to main dataset

Add 1st changepoint data to dataset: 

```{r}
canyonSites_cp <- merge(canyonSites_complete, cps)
```


Metrics of interest for each year are: 

  - Changepoint day of year (already have)
  - Peak runoff value & date
  - Total runoff volume from CP DOY thru 10/1 *NOTE* - need to set threshold for # missing days before doing this calc. 
  
Work flow: 

  - Extract data for CP1 thru 10/1 for each year
  - Calcualte peak runoff value and date 
  - Run models for these three metrics
  - Drop years w/ >"X" missing days of data
  - Calculate total runoff volume, plot over time.

##### c. Peak runoff value/Peak runoff DOY/Runoff start DOY Analysis

Extract data for runoff start to 10/1, calculate max discharge value and DOY, extract runoff start DOY
  
```{r}
canyonMetrics <- canyonSites_cp%>%
  group_by(siteYr)%>% #group by site/year
  filter(!doy<cpDoy, !doy>(doyMax-92))%>% #drop data from before CP day of year and after Sep 30th (NOT including 10/1 currently)
  ungroup()%>%
  group_by(site, year)%>% #group by site and year
  summarise(runoffStart = unique(cpDoy), 
            runoffPeak = max(discharge_cfs, na.rm = T), 
            peakDoy = doy[which.max(na.omit(discharge_cfs))])%>% #extract runoff start DOY, max discharge value, max discharge day of year for each site/year
  pivot_longer(!c(site, year), names_to = "metricName", values_to = "metricValue")
```

Nest data:


```{r}
canyonNest <- canyonMetrics %>%
  nest(data = -c(site, metricName)) #create new nested dataframe with columns for site and metricName, one column for data - data entries for each site include year and metricValues. 
```

Run basic models for each site and metric of interest: 

```{r}
canyonModels <- canyonNest %>%
  mutate(model = map(data, ~lm(metricValue~year, data = .)%>% #run a model of temp~year for each site (referenced by "."); store results in new column called "model"
                     tidy))%>% #tidy function to construct a summary table with model results (seperate tables for each site)
  unnest(model) %>% #unnest model column to get rows for each site, columns for model metrics
  filter(term == "year")  #remove intercept term to just get info on year for each site
canyonModels
```

Add P-values for plotting: 

```{r}
canyonPV <- merge(canyonMetrics, 
                     select(canyonModels, site, metricName, p.value) #add p-values
                     )%>%
  mutate(metricPrint = case_when(
    metricName == "runoffStart" ~ "Runoff start day of year", 
    metricName == "runoffPeak" ~"Peak discharge, cfs", 
    metricName == "peakDoy" ~"Peak discharge day of year" #add print-friendly metric name for plotting
  ))%>%
  mutate(siteMetric = paste(site, metricPrint, sep = " - ")) #add metric + site column for plotting

cyPVs <- canyonPV%>%
  group_by(siteMetric, metricPrint)%>%
  summarize(x = mean(year), label = round(unique(p.value), 2)) #create separate dataframe with p-values for adding to plots
```

Plotting:

```{r}
canyon.plot <- ggplot(canyonPV, aes(x = year, y = metricValue, color = metricPrint))+
  geom_point()+
  geom_smooth(se = T, method = "lm", linetype = "dashed")+
  theme_classic()+
  labs(x = "Year", y = "Metric Value", color = "Metric")+
  scale_color_manual(values = c("darkgreen","cyan4", "darkblue"))+
  geom_text(data = cyPVs, aes(x = x, label = paste("P=", label)), y = Inf, vjust = 2)+
  facet_wrap(~siteMetric, scales = "free", labeller = label_wrap_gen(width = 40))+
  theme(axis.text = element_text(size = 11), 
        axis.title = element_text(size = 11, face = "bold"), 
        strip.text = element_text(size = 10, face = "bold"), 
        legend.position = "bottom")

canyon.plot

#Save plot:
ggsave("CanyonSite_metrics.jpg", canyon.plot, device = "jpeg", path = here("outputs", "plots", "updatedTrends"), units = "in", dpi = "retina", height = 10, width = 10)
```


##### d. Runoff volume analysis

First, drop years with >5% of data missing and calculate runoff volumes:

```{r}
runoff <- canyonSites_cp%>%
  mutate(endDate = ymd(paste(year, "-09-30", sep = "")))%>% #add end date for later date re-filling
  group_by(siteYr, site, year)%>% #group by site/year
  filter(!doy<cpDoy, !doy>(doyMax-92))%>% #drop data from before CP day of year and after Sep 30th (NOT including 10/1 currently)
  complete(date = seq.Date(min(date), max(endDate), by = "day"))%>% #add missing days back (removed to calculate CP's), fill with NAs
  mutate(nMissing = length(which(is.na(discharge_cfs))), 
         nDays = yday(max(endDate, na.rm = T))-max(cpDoy, na.rm = T))%>% #count # NA's per runoff period and total # days per runoff period
  mutate(percMissing = nMissing/nDays)%>% #calculate % of data missing (# missing days/total days in runoff period)
  ungroup()%>%
  filter(!percMissing>0.05)%>% #drop years with >5% of data missing
  mutate(dischargeCf = discharge_cfs*60*12)%>% #calculate total cubic feet going by gauge for each day
  group_by(site, year)%>%
  summarize(runoffVol = sum(dischargeCf, na.rm = T), 
            runoffDays = unique(nDays)) #calculate total runoff volume for each year/site, runoff duration
```

Nest data:


```{r}
runoffNest <- runoff %>%
  nest(data = -c(site)) #create new nested dataframe with columns for site and metricName, one column for data - data entries for each site include year and metricValues. 
```

Run basic models for each site and metric of interest: 

```{r}
runoffModels <- runoffNest %>%
  mutate(model = map(data, ~lm(runoffVol~year, data = .)%>% #run a model of temp~year for each site (referenced by "."); store results in new column called "model"
                     tidy))%>% #tidy function to construct a summary table with model results (seperate tables for each site)
  unnest(model) %>% #unnest model column to get rows for each site, columns for model metrics
  filter(term == "year")  #remove intercept term to just get info on year for each site
runoffModels
```

Add P-values for plotting: 

```{r}
runoffPV <- merge(runoff, 
                     select(runoffModels, site, p.value) #add p-values
                     )

rfPVs <- runoffPV%>%
  group_by(site)%>%
  summarize(x = mean(year), label = round(unique(p.value), 2)) #create separate dataframe with p-values for adding to plots
```

Plotting:

```{r}
runoff.plot <- ggplot(runoffPV, aes(x = year, y = runoffVol, color = site))+
  geom_point()+
  geom_smooth(se = T, method = "lm", linetype = "dashed")+
  theme_classic()+
  labs(x = "Year", y = "Total runoff volume, cubic feet")+
  scale_color_manual(values = brewer.pal(length(runoffPV$site), "PRGn"))+
  geom_text(data = rfPVs, aes(x = x, label = paste("P=", label)), y = Inf, vjust = 2)+
  facet_wrap(~site, scales = "free", labeller = label_wrap_gen(width = 40))+
  theme(axis.text = element_text(size = 11), 
        axis.title = element_text(size = 11, face = "bold"), 
        strip.text = element_text(size = 10, face = "bold"), 
        legend.position = "none")

runoff.plot

#Save plot:
ggsave("Runoff_metrics.jpg", runoff.plot, device = "jpeg", path = here("outputs", "plots", "updatedTrends"), units = "in", dpi = "retina", height = 10, width = 10)
```

#### iii. Summer stream temperature

Data prep:

```{r}
canyonTemps <- fullData_usgs%>%
  filter(site == "Logan River near Franklin Basin"|site == "Wood Camp Spring"|site == "Logan River at First Dam"|site == "Ricks Spring above confluence with Logan River"|site == "Logan River at Guinavah Campground Bridge"|site == "Logan River near Tony Grove")%>% #extract canyon sites
  filter(parameter == "Water Temperature")%>% #extract temperature data
  select(!c(parameter, unit))%>% #drop extra cols
  rename(waterTemp = parameterValue)%>% #renaming cols
  mutate(year = year(dateTime), 
         month = month(dateTime), 
         doy = yday(dateTime), 
         date = ymd(substr(dateTime, 1, 10)))%>% #extract year, month, and date
  mutate(waterYear=ifelse(month<10, year, year+1))%>% #add water year column
  group_by(site,year)%>% 
  mutate(doyMin = min(doy), 
         doyMax = max(doy))%>% #calculate min and max DOY for each site/year
  mutate(waterDOY = ifelse(doy>=doyMax-91, 
                           doy-(doyMax-92), 
                           doy+92)
         )%>% #use min/max Doy to calculate water day of year
  ungroup()%>%
  filter(month==8) #extract August
```

Drop years with >5% missing data: 

```{r}
canyonTempsClean <- canyonTemps%>%
  group_by(site, year)%>% #group by site and year
  mutate(nMissing = length(which(is.na(waterTemp))), 
         nTot = n())%>% #calculate # missing obs, total # obs
  mutate(percMissing = (nMissing/nTot)*100)%>% #calculate % missing obs
  ungroup()%>%
  filter(!percMissing>5) #drop years with >5% data missing. 
```

Calculate metrics of interest; rearrange for plotting:

For BCT Temps: using 17 C based on this reference: https://ndep.nv.gov/uploads/water-wqs-docs/BonnevilleLCTTTA.pdf

```{r}
canyonTempMetrics <- canyonTempsClean %>%
  group_by(site, year, doy)%>%
  mutate(dailyMax = max(waterTemp, na.rm = T), 
         dailyMean = mean(waterTemp, na.rm = T))%>%
  ungroup()%>%
  mutate(dailyMax = ifelse(is.infinite(dailyMax), NA, dailyMax), 
         dailyMean = ifelse(is.infinite(dailyMean), NA, dailyMean))%>%
  group_by(site, year)%>%
  summarise(
    meanTemp = mean(waterTemp, na.rm = T), 
    meanMax = mean(dailyMax, na.rm = T), 
    medTemp = median(waterTemp, na.rm = T)
    )%>%
  pivot_longer(!c(site, year), names_to = "metric", values_to = "metricValue")%>%
  mutate(siteMetric = paste(site, metric, sep = " - "))%>%
  mutate(metricPrint = case_when(
    metric == "meanTemp" ~ "Mean August Stream Temp., C", 
    metric == "meanMax" ~ "Mean daily maximum August Stream Temp, C", 
    metric == "medTemp" ~ "Median August Stream Temp., C"
  ))
```

Run analysis: 

Nest data:


```{r}
CanyonTempNest <- canyonTempMetrics %>%
  nest(data = -c(siteMetric)) #create new nested dataframe with columns for site and metricName, one column for data - data entries for each site include year and metricValues. 
```

Run basic models for each site and metric of interest: 

```{r}
CanyonTempModels <- CanyonTempNest %>%
  mutate(model = map(data, ~lm(metricValue~year, data = .)%>% #run a model of temp~year for each site (referenced by "."); store results in new column called "model"
                     tidy))%>% #tidy function to construct a summary table with model results (seperate tables for each site)
  unnest(model) %>% #unnest model column to get rows for each site, columns for model metrics
  filter(term == "year")  #remove intercept term to just get info on year for each site
CanyonTempModels
```

Add P-values for plotting: 

```{r}
CanyonTempPV <- merge(canyonTempMetrics, 
                     select(CanyonTempModels, siteMetric, p.value) #add p-values
                     )%>%
  group_by(site, metric)%>%
  mutate(meanYear = mean(year))%>%
  ungroup()

ctPVs <- CanyonTempPV%>%
  group_by(siteMetric, metricPrint)%>%
  summarize(x = mean(year), label = round(unique(p.value), 2), y = Inf) #create separate dataframe with p-values for adding to plots
```

Plotting:

```{r}
canyon.temp.plot <- ggplot(CanyonTempPV, aes(x = year, y = metricValue, color = factor(metricPrint)))+
  geom_point()+
  geom_smooth(se = T, method = "lm", linetype = "dashed")+
  theme_classic()+
  labs(x = "Year", y = "Metric Value", color = "Legend")+
  scale_color_manual(values = c("orange", "darkred", "darkorange"))+
  #geom_hline(aes(yintercept = 17), color = "darkgrey", linetype = "dashed")+
  geom_text(data = ctPVs, aes(x = x, y = y,label = paste("P=", label)), vjust = 2)+
  facet_wrap(~siteMetric, scales = "free", labeller = label_wrap_gen(width = 40))+
  #geom_text(aes(x = meanYear, y = 17.1, label = "BCT Chronic Temp. Threshold"), color = "darkgrey", size = 2.5 )+
  theme(axis.text = element_text(size = 11), 
        axis.title = element_text(size = 11, face = "bold"), 
        strip.text = element_text(size = 10, face = "bold"), 
        legend.position = "bottom")

canyon.temp.plot

#Save plot:
ggsave("Canyon_summerTemps.jpg", canyon.temp.plot, device = "jpeg", path = here("outputs", "plots", "updatedTrends"), units = "in", dpi = "retina", height = 10, width = 10)
```


### D. Valley Sites - Discharge + Temperature Metrics

Valley sites currently included are: 

  - Water lab
  - Main St Bridge
  - Mendon Rd. 
  
For each site, metrics of interest are: 

  - Total water year discharge
  - Change in water year discharge between sites
  - Discharge vs. Water Temperature 
  
#### i. Discharge Metrics

##### a. Data prep & QA/QC

First, need to extract Valley sites and drop any water years with significant missing data. 

```{r}
valleySites <- fullData_usgs%>%
  filter(site=="Logan River at Main Street (Highway 89/91) Bridge"|site=="Logan River at the Utah Water Research Laboratory west bridge"|site=="Logan River at Mendon Road (600 South)")%>%
  mutate(year = year(dateTime), 
         month = month(dateTime), 
         date = ymd(substr(dateTime, 1, 10)), 
         doy = yday(dateTime))%>% #add year, month, date, and day of year columns
  mutate(waterYear=ifelse(month<10, year, year+1))%>%  #Calculate water year column.
  group_by(site,parameter, year)%>% #group by site, parameter, and year
  mutate(doyMin = min(doy), 
         doyMax = max(doy))%>% #calculate min and max DOY for each site/year
  mutate(waterDOY = ifelse(doy>=doyMax-91, 
                           doy-(doyMax-92), 
                           doy+92)
         )%>% #use min/max Doy to calculate water day of year 
  ungroup()%>%
  group_by(site, parameter, waterYear)%>% #group by site, parameter, and water year
  mutate(waterDoyMin = min(waterDOY), 
         waterDoyMax = max(waterDOY))%>% #Calculate min and max water day of year
  ungroup()%>%
  filter(!waterDoyMin>1, !waterDoyMax<365)%>% #drop years that start after water DOY 1 or end before water DOY 365
  group_by(site, parameter, waterYear)%>%
  mutate(nMissing = length(which(is.na(parameterValue))))%>%
  mutate(percMissing = (nMissing/n())*100)%>% #calculate number of missing obs and percentage of missing obs per site/parameter/year
  ungroup()

#Quick summary of missing data: 

valleyDataAv <- valleySites%>%
  group_by(site, parameter, waterYear)%>%
  summarise(missing = unique(nMissing), percMis = unique(percMissing))

View(valleyDataAv)
```

Good data availability, but still some years with substantial missing data.


######1. Using Regression to fill data gaps: 

Goal: replace data gaps by regressing between two endpoints. 

Workflow: 
  - Is data gap >5 days? If YES, leave as is. If NO:
  - What are the last measurements before the gap and the first measurements after the gap? 
  - Use these to calculate a linear formula (y = mx+b) for the missing period
  - Replace missing values based on formula calculated above: plug in x-value, use calculated m & b values to calculate y-value. 
  - Add a flag for the total # of measurements replaced by regression. 

Identify data gap lengths (takes a while to run with full valley dataset): 

```{r}
naLengths <- valleySites %>%
  mutate(isNa = is.na(parameterValue))%>% #new column with NAs flagged
  group_by(site, parameter)%>% #group by site and parameter
  mutate(naGrp = consecutive_id(isNa))%>% #add consecutive ID's for each group of NA's or non-NA's within each site/parameter
  mutate(naGrp = paste(site, parameter, naGrp, sep = "-"))%>% #add site and parameter to NA group label to make unique
  ungroup()%>% #ungroup
  group_by(naGrp)%>% #group by NA group
  mutate(nObs = n())%>% #calculate # of obs in NA group
  mutate(timeLength = (max(dateTime)-min(dateTime)))%>%
  mutate(nDays = as.numeric(timeLength)/(60*60*24))%>% #calculate length of group, convert to days
  ungroup()%>% #ungroup
  mutate(overFive = ifelse(nDays>5&isNa==T, T, F))%>% #add flag for NA groups over 5 days long
  mutate(obsNum = row_number()) #add row #
```

Next, extract values on either side of NA gaps and regress between them. 

Data Set Up: 

```{r}
naGroups <- naLengths %>%
  group_by(naGrp)%>% #group by NA Group
  mutate(minObs = ifelse(isNa == F, min(obsNum), min(obsNum-1)), 
         maxObs = ifelse(isNa == F, max(obsNum), max(obsNum+1)))%>% #extract minimum and maximum observation # for each group, add/subtract one for groups with NA's to get measurements on either end. 
  mutate(minObs20 = ifelse(naGrp == min(naGrp), minObs, minObs-20), 
         maxObs20 = ifelse(naGrp == max(naGrp), maxObs, maxObs+20))%>%
  ungroup()%>%
  #mutate(start = mean(c(parameterValue[minObs20:minObs]), na.rm = T), # Calculate mean value for 20 obs before gap
       #  end = mean(c(parameterValue[maxObs:maxObs20]), na.rm = T))%>% #Calculate mean value for 20 obs after gap
  mutate(start = parameterValue[minObs], 
         end = parameterValue[maxObs])%>%
  select(site, isNa, naGrp, minObs, maxObs, start, end)%>% #drop extra columns
  distinct()%>% #drop duplicate rows
  pivot_longer(!c(site, isNa, naGrp, start, end), names_to = "type", values_to = "obsNum")%>%
  pivot_longer(!c(site, isNa, naGrp, type, obsNum), names_to = "bound", values_to = "obsVal")%>% #pivot longer twice to get observation #'s and values into columns for regression
  filter((type =="minObs"&bound == "start")|(type == "maxObs"&bound=="end")) #Extract "matching" columns that have min obs # + starting value; max obs # and end value
```

Regress with for-loop: 

```{r}
coef<-data.frame() #empty dataframe to recieve data
na.groups <- unique(naGroups$naGrp) #vector of all NA groups

for(i in 1:length(na.groups)){
  dat <- filter(naGroups, naGrp == na.groups[i])
  
  mod <- lm(obsVal~obsNum, data = dat)
  b<-mod$coefficients[1]
  m<-mod$coefficients[2]
  
  df<-data.frame(na.groups[i], m, b, row.names = NULL)%>%
    rename(naGrp = 1, slope = 2, yInt = b)
  
  coef <- rbind(coef, df)
  
  print(paste("Finished group # ", na.groups[i], sep = ""))
}
```

Add back to primary dataframe, calculate missing values:

```{r}
regressed <- merge(naLengths, coef)%>% #add coefficients to naLengths dataframe
  mutate(paramValueReg = ifelse(isNa == FALSE|overFive==TRUE, #if NOT an NA group OR is an NA group with >5 day data gap:
                                parameterValue, #leave parameter value as is
                                (slope*obsNum)+yInt #if IS an NA group with <5 day datagap, use coefficients to replace missing values. 
                                ))%>%
  group_by(site, waterYear, parameter)%>% #group by site, water year, parameter
  mutate(nReplaced = nMissing - length(which(is.na(paramValueReg))))%>%
  ungroup() #calculate # of observations replaced and ungroup

#check # replaced per water year:
summary<-regressed%>%
  select(site, parameter, waterYear, nMissing, nReplaced)%>% #extract water year, site, parameter, # missing and # replaced, drop duplicate rows. 
  distinct()
```


Re-calculate data availability, clean up for analysis: 

```{r}
regFinal <- regressed%>%
  group_by(site, waterYear, parameter)%>% #group by site, water year parameter
  mutate(nMissingReg = length(which(is.na(paramValueReg))))%>%
  mutate(percMissingReg = (nMissingReg/n())*100, 
         percReplaced = (nReplaced/n())*100)%>% #calculate # missing after regresssion, % missing after regression
  select(site, parameter, unit, dateTime, parameterValue, paramValueReg, year, month, date, doy, waterYear, doyMin, doyMax, waterDOY, waterDoyMin, waterDoyMax, nMissingReg, percMissingReg, nReplaced, percReplaced)%>% #drop extra columns
  ungroup()%>%
  group_by(site, parameter)%>%
  mutate(totMissing = length(which(is.na(parameterValue))), 
         totMissingReg = length(which(is.na(paramValueReg))))%>%
  mutate(totReplaced = totMissing-totMissingReg)%>%
  mutate(totPercReplaced = (totReplaced/n())*100)%>%
  ungroup()%>%
  mutate(isRegressed = ifelse(is.na(parameterValue)==T&is.na(paramValueReg)==F, "Y", "N"))

unique(regFinal$isRegressed)

#re-check data availability:
regDataAv <- regFinal%>%
  group_by(site, parameter, waterYear)%>%
  summarise(missing = unique(nMissingReg), percMis = unique(percMissingReg), replaced = unique(nReplaced), percReplaced = unique(percReplaced))

View(regDataAv)
```


Visualization of regressed data:

```{r}
reg.plot <- ggplot(filter(
  regFinal, site == "Logan River at Main Street (Highway 89/91) Bridge"&parameter == "Discharge_cfs"&year==2022&month>6
), aes(x = dateTime, y = paramValueReg, color = isRegressed))+
  geom_point(size = 0.5)+
  scale_color_manual(values =c("black", "red"))+
  theme_classic()
reg.plot
```

#######1A. Alternative regression method

Goal = regress over more points to improve accuracy. For now - trying to regress over 20 points before and after gap. 

General workflow: 

  - Write function to iterate the following steps over each site: 
    + Calculate lengths of each sequence of NA and non-NA values
    + Extract the 20 observations before and after each gap
    + Regress over these 40 observations
    + Predict values for all NA values 
    + Add these predicted values back into dataframe
  - Apply function to valleySites data
  - Calculate % data regressed in each site

Write function to carry out steps above:

```{r}
##########################################################################
# fill_gaps_weighted()
#
# Fills gaps in observationValue using weighted linear regression
# (inverse squared distance weighting) separately for each site.
#
# Inputs:
#   df                 = data frame
#   siteColumn         = name of site/group column
#   observationColumn  = name of observation number column
#   valueColumn        = name of value column
#   n                  = number of non-missing observations to use
#                        before and after each gap (default = 20)
#
# Returns:
#   filledData         = original data with gaps filled
#   regressions        = regression information for each gap
#   skippedGaps        = gaps that could not be filled
##########################################################################

fill_gaps_weighted <- function(
    df,
    siteColumn = "Site",
    observationColumn = "observationNumber",
    valueColumn = "observationValue",
    n = 20 #defining input variables
){

  ############################################################
  # Make a copy of the original data
  ############################################################

  filled_df <- df

  ############################################################
  # Create empty tables for output
  ############################################################

  regression_summary <- data.frame()

  skipped_gaps <- data.frame()

  ############################################################
  # Process each site independently
  ############################################################

  for(site in unique(df[[siteColumn]])){

    ##########################################################
    # Extract one site's data
    ##########################################################

    site_rows <- which(df[[siteColumn]] == site) 

    site_df <- df[site_rows, ]

    ##########################################################
    # Sort observations by observation number
    ##########################################################

    site_df <- site_df[order(site_df[[observationColumn]]), ]

    ##########################################################
    # Identify runs of missing values
    ##########################################################

    na <- is.na(site_df[[valueColumn]]) #vector with TRUE for all NA values, FALSE for all non-NA values

    r <- rle(na) #calculate length of each sequence in vector created above

    ends <- cumsum(r$lengths) #extract row numbers for end of each sequence

    starts <- ends - r$lengths + 1 #extract row numbers for beginning of each sequence

    gap_starts <- starts[r$values] #Extract row # for start of each group of NA's

    gap_ends <- ends[r$values] #extract row # for end of each group of NA's
    

    ##########################################################
    # Process each gap
    ##########################################################

    for(i in seq_along(gap_starts)){

      gs <- gap_starts[i]
      ge <- gap_ends[i] #Extract start and end row # for "ith" gap
      
      

      ########################################################
      # Find the last n non-missing observations before gap
      ########################################################

      before_idx <- which(
        !is.na(site_df[[valueColumn]][1:(gs-1)]) #extract all non NA values before gap
      )

      before_idx <- tail(before_idx, n) #then extract last 20 values

      ########################################################
      # Find the first n non-missing observations after gap
      ########################################################

      after_idx <- which(
        !is.na(site_df[[valueColumn]][(ge+1):nrow(site_df)]) #extract all non-NA values after gap
      )

      if(length(after_idx) > 0)
        after_idx <- head(after_idx, n) + ge #if there are non-NA values after the gap, extract the first 20. 

      ########################################################
      # Skip if there are not enough observations
      ########################################################

      if(length(before_idx) < n ||
         length(after_idx) < n){ #if there are <20 non-NA values before or after gap, skip 

        skipped_gaps <- rbind(
          skipped_gaps,
          data.frame(
            Site = site,
            gapStart = gs,
            gapEnd = ge,
            observationsBefore = length(before_idx),
            observationsAfter = length(after_idx)
          ) #and print gap info in skipped_gaps table
        )

        next

      }

      ########################################################
      # Build regression data
      ########################################################

      fit_idx <- c(before_idx, after_idx) #vector w. row numbers for 20 obs before and after gap

      fit_data <- data.frame( 

        x = site_df[[observationColumn]][fit_idx],

        y = site_df[[valueColumn]][fit_idx] #extract observation values and numbers for 20 obs before and after gap

      )

      ########################################################
      # Calculate distance from each observation to the gap
      ########################################################

      distance <- ifelse(

        fit_idx < gs,

        gs - fit_idx,

        fit_idx - ge

      ) #calculate distance from gap for each obs - obs adjacent to gap is distance 1, next obs is distance 2, etc. 

      ########################################################
      # Inverse squared distance weights
      ########################################################

      weights <- 1 / distance^2 #calculate weight = 1 / distance squared (closest values get highest weight)

      ########################################################
      # Fit weighted linear regression
      ########################################################

      fit <- lm(

        y ~ x,

        data = fit_data #,

       # weights = weights

      ) #fit model using weights

      ########################################################
      # Predict missing values
      ########################################################

      gap_idx <- gs:ge #ID gap row numbers

      predictions <- predict(

        fit,

        newdata = data.frame(

          x = site_df[[observationColumn]][gap_idx]

        ) # then predict values for these row numbers using the coefficients calculated in the LM above

      )

      ########################################################
      # Replace missing values
      ########################################################

      site_df[[valueColumn]][gap_idx] <- predictions #replace missing values in site_df with predicted values

      ########################################################
      # Save regression statistics
      ########################################################

      regression_summary <- rbind(

        regression_summary,

        data.frame(

          Site = site,

          gapStart = gs,

          gapEnd = ge,

          gapLength = ge - gs + 1,

          nPoints = nrow(fit_data),

          slope = coef(fit)[2],

          intercept = coef(fit)[1],

          rSquared = summary(fit)$r.squared

        )

      )

    } 

    ##########################################################
    # Put this site's completed data back into the full dataset
    ##########################################################

    filled_df[site_rows, ] <- site_df

  } 

  ############################################################
  # Return results
  ############################################################

  return(

    list(

      filledData = filled_df,

      regressions = regression_summary,

      skippedGaps = skipped_gaps

    )

  )

}
```

Prep valleySites data to go into function

```{r}
valleySimple <- valleySites%>%
  select(site, parameter, dateTime, parameterValue)%>% #drop extra cols
  mutate(siteParam = paste(site, parameter, sep = " - "))%>% #add SiteParameter column for grouping
  group_by(siteParam)%>%
  mutate(obsNum = row_number()) #add observation # for each site/parameter group
```

```{r}
regOutput <- fill_gaps_weighted(valleySimple, 
                   siteColumn = "siteParam", 
                   observationColumn = "obsNum", 
                   valueColumn = "parameterValue",
                   n = 20
                   )

regOutput$regressions
summary(regOutput$regressions$rSquared)
```

Identify regressed values, draw plots to check accuracy. 

```{r}
valleyReg <- regOutput$filledData%>% #extract filled dataframe from function output
  rename(paramReg = parameterValue)%>% #renaming column with regressed data
  merge(valleySimple)%>% #add back to original data
  mutate(isReg = ifelse(is.na(parameterValue)==T&is.na(paramReg)==F, "Y", "N"))%>%
  mutate(year = year(dateTime), 
         month = month(dateTime))%>%
  group_by(siteParam)%>%
  mutate(isNa = is.na(parameterValue))%>% #add column with T/F for NAs
  mutate(naGrp = consecutive_id(isNa))%>% #add consecutive ID's for each group of NA's or non-NA's within each site/parameter
  ungroup()

valleyReg[2950:2960,]

naGrps <- valleyReg %>%
  filter(isNa == T)%>%
  select(siteParam, naGrp)%>%
  distinct()

naGrps
```

Visualization of regressed data:

```{r}
reg.plot2 <- ggplot(filter(
  valleyReg, site == "Logan River at Main Street (Highway 89/91) Bridge"&parameter == "Discharge_cfs"&year==2022&month>6
), aes(x = dateTime, y = paramReg, color = isReg))+
  geom_point(size = 0.5)+
  scale_color_manual(values =c("black", "red"))+
  theme_classic()
reg.plot2
```

Seems to be working well - needs some tuning to:

  - Don't fill gaps if >5 days long
  - Maybe don't fill gaps with bad R2... <0.2? other thresholds?
  
First - need to associate regression info with each gap to decide whether or not to include it. 

```{r}
regStats <- regOutput$regressions%>%
  group_by(Site)%>% #group by site
  mutate(naGrp = 2*row_number())%>% #add NA group for merge
  ungroup()%>%
  rename(siteParam = Site)
regStats
```

Adding regression summary data to valleyReg dataframe:

```{r}
valleyRegStat <- merge(valleyReg, regStats, all = T)%>% #merge
  arrange(siteParam, dateTime) #sort by date time and site parameter

head(valleyRegStat) #Check beginning of data frame

valleyRegStat[2950:2960,] #check 1st data gap
valleyRegStat[278662:278680,] #check random longer data gap
```

Apply filters and calculate data availability - Currently not filling gaps with: 

  - >480 missing obs (= to 5 missing days @ 1 obs every 15min)
  - R2 <0.5
  
```{r}
valleyRegCor <- valleyRegStat %>%
  mutate(waterYear = ifelse(month<10, year, year+1))%>%
  mutate(paramReg = ifelse(isNa ==T&(gapLength>480|rSquared<0.5), NA, paramReg))%>% #for NA groups: change paramReg back to NA if gap length >480 OR R2 <0.5
  mutate(isReg = ifelse(is.na(parameterValue)==T&is.na(paramReg)==F, "Y", "N"))%>% #re-calculate whether point was regressed after filters
  group_by(siteParam, waterYear)%>%
  mutate(nObs = n(), 
         nReg = length(which(isReg == "Y")), 
         nMissing = length(which(is.na(paramReg))))%>%
  mutate(percMissing = (nMissing/nObs)*100, 
         percReg = (nReg/nObs)*100)%>% #calculate # obs, # observations regressed, # missing obs, % missing obs, and % regressed obs for each year.
  ungroup()
  


valleyRegCor[2950:2960,] #check 1st data gap
valleyRegCor[278662:278680,] #check random longer data gap
```

-----
-----

#STOPPING HERE ON 7/23 - need to re-code all valley analyses after this point using updated regression data. Idea for next time - write a function that will calculate % missing and % regressed for a subset of the data - will be doing this a bunch for lots of different time periods, would be nice to not have to code out every time. 

######2. Re-check data availability and prep for analysis. 

Drop years with >5% missing, calculate discharge metrics:

```{r}
valleyDischarge <- valleyRegCor%>%
  filter(parameter=="Discharge_cfs", !percMissing>5)%>% #drop years with >5% missing data after regression
  select(!c(parameter))%>% 
  rename(dischargeCfs = paramReg)%>% #dropping xtra cols, renaming
  mutate(dischargeCf = dischargeCfs*15)%>% #calculate total cubic feet going by gauge for each 15sec interval
  group_by(site)%>%
  mutate(totPercReplaced = (sum(nReg)/sum(nObs))*100)%>%
  ungroup()%>%
  group_by(site, waterYear)%>% 
  summarise(totDischarge = sum(dischargeCf, na.rm = T), 
            nReplaced = unique(nReg), 
            percReplaced = unique(percReg), 
            totPercReplaced = unique(totPercReplaced)
            ) #calculate total water year discharge for each site; include columns for labeling % replaced data for each site/year
```

##### b. Total Annual Discharge analysis

Nest data:


```{r}
valleyNest <- valleyDischarge %>%
  nest(data = -c(site)) #create new nested dataframe with columns for site and metricName, one column for data - data entries for each site include year and metricValues. 
```

Run basic models for each site and metric of interest: 

```{r}
valleyModels <- valleyNest %>%
  mutate(model = map(data, ~lm(totDischarge~waterYear, data = .)%>% #run a model of temp~year for each site (referenced by "."); store results in new column called "model"
                     tidy))%>% #tidy function to construct a summary table with model results (seperate tables for each site)
  unnest(model) %>% #unnest model column to get rows for each site, columns for model metrics
  filter(term == "waterYear")  #remove intercept term to just get info on year for each site
valleyModels
```

Add P-values for plotting: 

```{r}
valleyPV <- merge(valleyDischarge, 
                     select(valleyModels, site, p.value) #add p-values
                     )

vaPVs <- valleyPV%>%
  group_by(site)%>%
  summarize(x = mean(waterYear), label = round(unique(p.value), 2), repLabel = round(unique(totPercReplaced), 2), yRep = min(totDischarge)*1.25) #create separate dataframe with p-values and amount data replaced for adding to plots
```

Plotting:

```{r}
valley.plot <- ggplot(valleyPV, aes(x = waterYear, y = totDischarge, color = site))+
  geom_point()+
  geom_smooth(se = T, method = "lm", linetype = "dashed")+
  theme_classic()+
  labs(x = "Water year", y = "Total water year discharge, cubic feet")+
  scale_color_manual(values = c("cyan3","cyan4", "lightblue"))+
  geom_text(data = vaPVs, aes(x = x, label = paste("P=", label)), y = Inf, vjust = 2)+
  geom_text(data = vaPVs, aes(x = x, y = yRep, label = paste("Incl. ", repLabel, "% regressed data", sep = "")), vjust = 2)+
  facet_wrap(~site, scales = "free", labeller = label_wrap_gen(width = 40))+
  theme(axis.text = element_text(size = 11), 
        axis.title = element_text(size = 11, face = "bold"), 
        strip.text = element_text(size = 10, face = "bold"), 
        legend.position = "none")

valley.plot

#Save plot:
ggsave("Valley_discharge.jpg", valley.plot, device = "jpeg", path = here("outputs", "plots", "updatedTrends"), units = "in", dpi = "retina", height = 6, width = 10)
```


##### c. Change in total annual discharge

Data Prep/ QA/AC: Extract data for irrigation season; calculate amount of missing obs during irrigation season, drop years with >5% missing data, then calculate diversion amounts:

```{r}
diversion <- regFinal%>%
  filter(doy>(doyMax-261), doy<(doyMax-91))%>% #extract  data for irrigation season (4/15-9/30) 
  filter(parameter == "Discharge_cfs")%>% #extract discharge data
  select(dateTime, site, waterYear, paramValueReg)%>% #drop extra cols
  pivot_wider(id_cols = c(dateTime, waterYear), names_from = site, values_from = paramValueReg)%>% #pivot wider to get columns for each site
  rename(mainSt = 3, mendon = 4, waterLab = 5)%>% #Renaming cols
  mutate(waterLab_mendon = waterLab-mendon, 
         waterLab_main = waterLab-mainSt, 
         main_mendon = mainSt-mendon)%>% #calculate differences in discharge between each site for each dateTime
  select(dateTime, waterYear, waterLab_main, waterLab_mendon, main_mendon)%>% #Dropping extra cols
  pivot_longer(-c(dateTime, waterYear), names_to = "sites", values_to = "dischargeDif")%>% #pivot longer to get back to analysis-friendly format
  group_by(sites, waterYear)%>% #group by site and water year
  mutate(nMissing = length(which(is.na(dischargeDif))))%>%
  mutate(percMissing = (nMissing/n())*100)%>% #re-calculate # missing and % missing obs
  filter(!percMissing>5)%>% #drop seasons with >3% missing data
  ungroup()%>%
  mutate(dischargeCf = dischargeDif*15)%>% #calculate total discharge (cubic feet) going past gauge for each 15sec interval
  group_by(sites, waterYear)%>% #group by site and water year
  summarise(totQDiff = sum(dischargeCf, na.rm = T)) #Calculate total difference per irrigation season for each site/water year

## Calculate % regressed data included in analysis: 

regSum <- valleyRegCor %>%
  mutate(date = ymd(substr(dateTime, 1, 10)))%>%
  mutate(doy = yday(date))%>%
  group_by(site, parameter, year)%>%
  mutate(doyMax = max(doy), doyMin = min(doy))%>%
  filter(doy>(doyMax-261), doy<(doyMax-91))%>% #extract  data for irrigation season (4/15-9/30) 
  filter(parameter == "Discharge_cfs")%>% #extract discharge data
  group_by(site)%>%
  summarise(
    nMissing = length(which(is.na(parameterValue))), 
    nMissingReg = length(which(is.na(paramReg))), 
    nTot = n()
  )%>%
  mutate(percReplaced = ((nMissing-nMissingReg)/nTot)*100)%>%
  select(site, percReplaced)%>%
  pivot_wider(names_from = site, values_from = percReplaced)%>%
  rename(mainSt = 1, mendon = 2, waterLab = 3)%>% #Renaming cols
  mutate(
    main_mendon = mainSt+mendon, 
    waterLab_main = mainSt+waterLab, 
    waterLab_mendon = mendon+waterLab
  )%>%
  select(3:6)%>%
  pivot_longer(!c(waterLab), names_to = "sites", values_to = "totPercReplaced")%>%
  select(!waterLab)

```

Run analysis: 

Nest data:


```{r}
diversionNest <- diversion%>%
  nest(data = -c(sites)) #create new nested dataframe with columns for site and metricName, one column for data - data entries for each site include year and metricValues. 
```

Run basic models for each site and metric of interest: 

```{r}
diversionModels <- diversionNest %>%
  mutate(model = map(data, ~lm(totQDiff~waterYear, data = .)%>% #run a model of temp~year for each site (referenced by "."); store results in new column called "model"
                     tidy))%>% #tidy function to construct a summary table with model results (seperate tables for each site)
  unnest(model) %>% #unnest model column to get rows for each site, columns for model metrics
  filter(term == "waterYear")  #remove intercept term to just get info on year for each site
diversionModels
```

Add P-values for plotting: 

```{r}
diversionPV <- merge(diversion, regSum)%>%
  merge(select(diversionModels, sites, p.value) #add p-values
                     )

divPVs <- diversionPV%>%
  group_by(sites)%>%
  summarize(x = mean(waterYear), label = round(unique(p.value), 2), repLabel = round(unique(totPercReplaced), 2), yRep = quantile(totQDiff, 0.05)) #create separate dataframe with p-values for adding to plots
```

Plotting:

```{r}
diversion.plot <- ggplot(diversionPV, aes(x = waterYear, y = totQDiff, color = sites))+
  geom_point()+
  geom_smooth(se = T, method = "lm", linetype = "dashed")+
  theme_classic()+
  labs(x = "Water year", y = "Difference in irrigation season discharge, cubic feet")+
  scale_color_manual(values = c("cyan3","cyan4", "lightblue"))+
  geom_text(data = divPVs, aes(x = x, label = paste("P=", label)), y = Inf, vjust = 2)+
  geom_text(data = divPVs, aes(x = x, y = yRep, label = paste("Incl. ", repLabel, "% regressed data", sep = "")), vjust = 2)+
  facet_wrap(~sites, scales = "free", labeller = label_wrap_gen(width = 40))+
  theme(axis.text = element_text(size = 11), 
        axis.title = element_text(size = 11, face = "bold"), 
        strip.text = element_text(size = 10, face = "bold"), 
        legend.position = "none")

diversion.plot

#Save plot:
ggsave("Valley_diversions.jpg", diversion.plot, device = "jpeg", path = here("outputs", "plots", "updatedTrends"), units = "in", dpi = "retina", height = 6, width = 10)
```

##### d. Diversion volume vs. Total Runoff volume

Data Prep: 

```{r}
diversionMerge <- diversion %>%
  ungroup()%>%
  filter(sites == "waterLab_mendon")%>% #extract data for main vs. Mendon
  select(waterYear, totQDiff) #drop extra cols

dischargeDiversion <- merge(valleyDischarge, diversionMerge)%>% #merge with discharge data
  arrange(site, waterYear)%>%
  filter(site == "Logan River at the Utah Water Research Laboratory west bridge")
```

Run models: 

Nest data:

```{r}
diversionQNest <- dischargeDiversion%>%
  nest(data = -c(site)) #create new nested dataframe with columns for site and metricName, one column for data - data entries for each site include year and metricValues. 
```

Run basic models for each site and metric of interest: 

```{r}
diversionQModels <- diversionQNest %>%
  mutate(model = map(data, ~lm(totQDiff~totDischarge, data = .)%>% #run a model of temp~year for each site (referenced by "."); store results in new column called "model"
                     tidy))%>% #tidy function to construct a summary table with model results (seperate tables for each site)
  unnest(model) %>% #unnest model column to get rows for each site, columns for model metrics
  filter(term == "totDischarge")  #remove intercept term to just get info on year for each site
diversionQModels
```

Add P-values for plotting: 

```{r}
diversionQPV <- dischargeDiversion%>%
  merge(select(diversionQModels, site, p.value) #add p-values
                     )

divQPVs <- diversionQPV%>%
  group_by(site)%>%
  summarize(x = mean(totDischarge), label = round(unique(p.value), 2), repLabel = round(unique(totPercReplaced), 2), yRep = quantile(totQDiff, 0.05)) #create separate dataframe with p-values for adding to plots
```

Plotting:

```{r}
diversion.q.plot <- ggplot(diversionQPV, aes(x = totDischarge, y = totQDiff, color = site))+
  geom_point()+
  geom_smooth(se = T, method = "lm", linetype = "dashed")+
  theme_classic()+
  labs(x = "Total water year discharge at Water Lab, cubic feet", y = "Difference in irrigation season discharge, Water Lab to Mendon, cubic feet")+
  scale_color_manual(values = c("cyan3","cyan4", "lightblue"))+
  geom_text(data = divQPVs, aes(x = x, label = paste("P=", label)), y = Inf, vjust = 2)+
  geom_text(data = divQPVs, aes(x = x, y = yRep, label = paste("Incl. ", repLabel, "% regressed data", sep = "")), vjust = 2)+
  #facet_wrap(~site, scales = "free", labeller = label_wrap_gen(width = 40))+
  theme(axis.text = element_text(size = 11), 
        axis.title = element_text(size = 11, face = "bold"), 
        strip.text = element_text(size = 10, face = "bold"), 
        legend.position = "none")

diversion.q.plot

#Save plot:
ggsave("Valley_diversion_discharge.jpg", diversion.q.plot, device = "jpeg", path = here("outputs", "plots", "updatedTrends"), units = "in", dpi = "retina", height = 6, width = 10)
```

##### d. Diversion volume as % of annual volume - Change over time

Data Prep: 

```{r}
dischargeDiversionProp <- dischargeDiversion%>%
  mutate(diversionProp = (totQDiff/totDischarge)*100)
```

Run models: 

Nest data:

```{r}
diversionPropNest <- dischargeDiversionProp%>%
  nest(data = -c(site)) #create new nested dataframe with columns for site and metricName, one column for data - data entries for each site include year and metricValues. 
```

Run basic models for each site and metric of interest: 

```{r}
diversionPropModels <- diversionPropNest %>%
  mutate(model = map(data, ~lm(diversionProp~waterYear, data = .)%>% #run a model of temp~year for each site (referenced by "."); store results in new column called "model"
                     tidy))%>% #tidy function to construct a summary table with model results (seperate tables for each site)
  unnest(model) %>% #unnest model column to get rows for each site, columns for model metrics
  filter(term == "waterYear")  #remove intercept term to just get info on year for each site
diversionPropModels
```

Add P-values for plotting: 

```{r}
diversionPropPV <- dischargeDiversionProp%>%
  merge(select(diversionPropModels, site, p.value) #add p-values
                     )

divPropPVs <- diversionPropPV%>%
  group_by(site)%>%
  summarize(x = mean(waterYear), label = round(unique(p.value), 2), repLabel = round(unique(totPercReplaced), 2), yRep = quantile(diversionProp, 0.05)) #create separate dataframe with p-values for adding to plots
```

Plotting:

```{r}
diversion.prop.plot <- ggplot(diversionPropPV, aes(x = waterYear, y = diversionProp, color = site))+
  geom_point()+
  geom_smooth(se = T, method = "lm", linetype = "dashed")+
  theme_classic()+
  labs(x = "Water Year", y = "% difference in discharge, Water Lab to Mendon, cubic feet")+
  scale_color_manual(values = c("cyan3","cyan4", "lightblue"))+
  geom_text(data = divPropPVs, aes(x = x, label = paste("P=", label)), y = Inf, vjust = 2)+
  geom_text(data = divPropPVs, aes(x = x, y = yRep, label = paste("Incl. ", repLabel, "% regressed data", sep = "")), vjust = 2)+
  #facet_wrap(~site, scales = "free", labeller = label_wrap_gen(width = 40))+
  theme(axis.text = element_text(size = 11), 
        axis.title = element_text(size = 11, face = "bold"), 
        strip.text = element_text(size = 10, face = "bold"), 
        legend.position = "none")

diversion.prop.plot

#Save plot:
ggsave("Valley_diversion_proportion.jpg", diversion.prop.plot, device = "jpeg", path = here("outputs", "plots", "updatedTrends"), units = "in", dpi = "retina", height = 6, width = 10)
```

### ii. Water Temperature Metrics

#### a. Data prep & QA/QC

First, extract years with both discharge and water temperature data available for August 1 - Sept 30; drop years with >5% missing temperature data. 

```{r}
valleyTemps <- valleyRegCor %>%
  mutate(date = ymd(substr(dateTime, 1, 10)))%>%
  mutate(doy = yday(date))%>%
  group_by(site, parameter, year)%>%
  mutate(doyMax = max(doy), doyMin = min(doy))%>%
  filter(!doyMax<doyMax-92, !doyMin>doyMax-152)%>% #drop years that start after 8/1 or end before 9/30
  filter(doy>=doyMax-152 & doy<=doyMax-92)%>% #from remaining data, extract data for 7/1-8/31
  select(dateTime, site, parameter, paramReg, year)%>% #drop extra columns
  group_by(site, parameter, year)%>%
  pivot_wider(id_cols = c(dateTime, site, year), names_from = parameter, values_from = paramReg)%>% #pivot wider to get cols for discharge and water temp
  rename(dischargeCfs = 4, waterTemp = 5)%>% #renaming
  group_by(site, year)%>% # group by site/year/parameter
  mutate(qMissing = length(which(is.na(dischargeCfs))),
         tMissing = length(which(is.na(waterTemp))),
         nObs = n())%>%
  mutate(qPercMissing = (qMissing/nObs)*100, 
         tPercMissing = (tMissing/nObs)*100)%>% # calculate number missing obs and percent missing obs for each parameter/year
  filter(!tPercMissing>5) #drop years with >5% missing temp data after regression

#calculate amount of regressed data for Aug 1 - Sept 30 after dropping years with >5% missing

valleyTempReg <- valleyRegCor %>%
  mutate(date = ymd(substr(dateTime, 1, 10)))%>%
  mutate(doy = yday(date))%>%
  group_by(site, parameter, year)%>%
  mutate(doyMax = max(doy), doyMin = min(doy))%>%
  filter(!doyMax<doyMax-92, !doyMin>doyMax-152)%>% #drop years that start after 7/1 or end before 8/31
  filter(doy>=doyMax-152 & doy<=doyMax-92)%>% #from remaining data, extract data for 7/1-8/31
  group_by(site, year, parameter)%>%
  mutate(percMissing = (length(which(is.na(paramReg)))/n())*100)%>% #calculate # NA's in regressed data during period of interest
  ungroup()%>%
  filter(!percMissing>5)%>%
  group_by(site, parameter)%>%
  summarise(nReplaced = length(which(is.na(parameterValue)))-length(which(is.na(paramReg))), 
            nTot = n())%>%
  mutate(percRegressed = (nReplaced/nTot)*100)%>%
  select(site, parameter, percRegressed)

```

#### b. Average summer temperature analysis

Calculate average summer temp:

```{r}
summerTemps <- valleyTemps %>%
  mutate(date = ymd(substr(dateTime, 1, 10)))%>%
  group_by(site, year, date)%>%
  mutate(dailyMax = max(waterTemp, na.rm = T))%>%
  ungroup()%>%
  group_by(site, year)%>%
  summarise(meanTemp = mean(waterTemp, na.rm = T), 
            maxXbar =mean(dailyMax, na.rm = T), 
            medTemp = median(waterTemp, na.rm = T))%>%
  pivot_longer(!c(site, year), names_to = "metric", values_to = "metricValue")%>%
  filter(!metricValue == -Inf)%>%
  mutate(metricPrint = case_when(metric == "meanTemp"~ "Mean Aug-Sep Stream Temp., C", 
                              metric =="maxXbar"~"Mean daily max Aug-Sep Stream Temp., C", 
                              metric == "medTemp" ~"Median Aug-Sep Stream Temp., C"))%>%
  mutate(siteMetric = paste(site, metricPrint, sep = " "))%>%
  merge(filter(valleyTempReg, parameter == "Water Temperature"))
```

Run analysis: 

Nest data:


```{r}
tempNest <- summerTemps %>%
  nest(data = -c(siteMetric)) #create new nested dataframe with columns for site and metricName, one column for data - data entries for each site include year and metricValues. 
```

Run basic models for each site and metric of interest: 

```{r}
tempModels <- tempNest %>%
  mutate(model = map(data, ~lm(metricValue~year, data = .)%>% #run a model of temp~year for each site (referenced by "."); store results in new column called "model"
                     tidy))%>% #tidy function to construct a summary table with model results (seperate tables for each site)
  unnest(model) %>% #unnest model column to get rows for each site, columns for model metrics
  filter(term == "year")  #remove intercept term to just get info on year for each site
tempModels
```

Add P-values for plotting: 

```{r}
tempPV <- merge(summerTemps, 
                     select(tempModels, siteMetric, p.value) #add p-values
                     )

tPVs <- tempPV%>%
  group_by(siteMetric, metricPrint)%>%
  summarize(x = mean(year), label = round(unique(p.value), 2), yRep = quantile(metricValue, 0.1), repLabel = round(unique(percRegressed), 2)) #create separate dataframe with p-values for adding to plots
```

Plotting:

```{r}
temp.plot <- ggplot(tempPV, aes(x = year, y = metricValue, color = factor(metricPrint)))+
  geom_point()+
  geom_smooth(se = T, method = "lm", linetype = "dashed")+
  theme_classic()+
  labs(x = "Year", y = "Metric Value", color = "Legend")+
  scale_color_manual(values = c("orange", "darkred", "darkorange"))+
  geom_text(data = tPVs, aes(x = x, label = paste("P=", label)), y = Inf, vjust = 2)+
  geom_text(data = tPVs, aes(x = x, y = yRep, label = paste("Incl. ", repLabel, "% regressed data", sep = "")), vjust = 2)+
  facet_wrap(~siteMetric, scales = "free", labeller = label_wrap_gen(width = 40))+
  theme(axis.text = element_text(size = 11), 
        axis.title = element_text(size = 11, face = "bold"), 
        strip.text = element_text(size = 10, face = "bold"), 
        legend.position = "bottom")

temp.plot

#Save plot:
ggsave("Valley_summerTemps.jpg", temp.plot, device = "jpeg", path = here("outputs", "plots", "updatedTrends"), units = "in", dpi = "retina", height = 10, width = 10)
```

#### c. Discharge vs. Water Temperature

Data prep - get rid of years with >5% discharge OR water temperature, nest by site

```{r}
summerQTNest <- valleyTemps%>%
  filter(!qPercMissing>5)%>% #drop years with >5% missing discharge data
  nest(data = -c(site))
```

Run basic models for each site and metric of interest: 

```{r}
tempqModels <- summerQTNest %>%
  mutate(model = map(data, ~lm(waterTemp~dischargeCfs, data = .)%>% #run a model of temp~year for each site (referenced by "."); store results in new column called "model"
                     tidy))%>% #tidy function to construct a summary table with model results (seperate tables for each site)
  unnest(model) %>% #unnest model column to get rows for each site, columns for model metrics
  filter(term == "dischargeCfs")  #remove intercept term to just get info on year for each site
tempqModels
```

Add P-values for plotting: 

```{r}
tempqPV <- merge(unnest(summerQTNest, cols = c(data)), 
                     select(tempqModels, site, p.value) #add p-values
                     )

tqPVs <- tempqPV%>%
  group_by(site)%>%
  summarize(x = mean(dischargeCfs, na.rm = T), label = round(unique(p.value), 2)) #create separate dataframe with p-values for adding to plots
```

Plotting:

```{r}
temp.q.plot <- ggplot(tempqPV, aes(x = dischargeCfs, y = waterTemp, color = site))+
  geom_point()+
  geom_smooth(se = T, method = "lm", linetype = "dashed")+
  theme_classic()+
  labs(x = "Discharge, cfs", y = "Water temperature, C")+
  scale_color_manual(values = c("cyan3","cyan4", "lightblue"))+
  geom_text(data = tqPVs, aes(x = x, label = paste("P=", label)), y = Inf, vjust = 2)+
  facet_wrap(~site, scales = "free", labeller = label_wrap_gen(width = 40))+
  theme(axis.text = element_text(size = 11), 
        axis.title = element_text(size = 11, face = "bold"), 
        strip.text = element_text(size = 10, face = "bold"), 
        legend.position = "none")

temp.q.plot

#Save plot:
ggsave("Valley_dischargeTemps.jpg", temp.q.plot, device = "jpeg", path = here("outputs", "plots", "updatedTrends"), units = "in", dpi = "retina", height = 6, width = 10)
```



