```{r setup, include=F, echo = F}
library(tidyverse)
library(broom)
library(ggplot2)
library(lubridate)
library(RColorBrewer)
library(here)
library(reticulate)
library(patchwork)
library(snotelr)

py_require("pip")
py_require("hydroserverpy")
py_require("datetime")


knitr::opts_knit$set(root.dir = here("outputs", "htmls"))
```

## Overview

This script performs initial data cleanining, visualization, and basic analysis for five test sites in the Logan River Observatory network. These sites include: 

- Logan River at Franklin Basin (discharge, water temp)
- Logan River at Main Street (discharge, water temp)
- Logan River at Mendon (discharge, water temp)
- Franklin Basin Climate Station (Precipitation, Air temp)
- Logan River Golf Course Climate Station (Precipitation, Air temp)

Data is called directly from the LRO Hydroserver API (https://lro.hydroserver.org/browse)

This script performs the following actions: 

1. Read in and clean data files
  + Call Metadata for all sites in the LRO network
  + Filter out desired sites
  + Import data using a for loop
2. Plotting and basic analysis
  + Discharge
  + Stream Temperature

## I. Import Data

### A. LRO Data

First, Import necessary packages; establish connection to Hydroserver

```{python}
import hydroserverpy
import datetime
import pandas as pd

from datetime import datetime

from hydroserverpy import HydroServer

#initialize Hydroserver Connection with login Credentials **NOTE** you may need to update these to your Hydroserver login info

hs_api = HydroServer(host = 'https://lro.hydroserver.org', email = 'ggianniny@gmail.com',password='syxqyz-Zeqpor-7naxki')

#Load info for all datastreams:

datastreams = hs_api.datastreams.list(fetch_all = True)

```

#### i. Call metadata for all "datastreams" (i.e. all parameters at all sites) in the LRO network

Setup empty data frame to recieve data:

```{python}
df = pd.DataFrame(data = None, index = None, columns = ['uid', 'name', 'property','propCode','unit', 'medium', 'procLevel']) #initialize empty dataframe (outside loop)
```

Read in data with a for-loop (NOTE - this will take a few minutes to run)

```{python}
for i in range(0, len(datastreams.items)):
        uid = str(datastreams.items[i].uid) #extract UID
        name = str(datastreams.items[i].thing.name) #extract site name
        prop = str(datastreams.items[i].observed_property.name) #extract measured property
        propCode = str(datastreams.items[i].observed_property.code) #extract measured property code
        unit = str(datastreams.items[i].unit.name) #extract unit
        medium = str(datastreams.items[i].sampled_medium) #extract sampled medium
        procLevel = str(datastreams.items[i].processing_level.definition) #extract processing level
        siteInfo = pd.DataFrame(data = {'uid':[uid], 'name':[name], 'property':[prop], 'propCode':[propCode], 'unit':[unit],'medium':[medium], 'procLevel':[procLevel]}) #convert uid, name, and unit into dataframe 
        df = pd.concat([df, siteInfo]) #add siteInfo to df
        print("working on"+" "+str(i)) #print current site for progress tracking
```

#### ii. Filter out desired sites/parameters

First - convert to python, filter out quality controlled data, create reference DF's

```{r}
r_df <- py_to_r(py$df)%>% #converting to R for easier wrangling
  filter(procLevel == "Quality controlled data"|procLevel == "Derived products")%>% #extract quality controlled and derived product data only
  arrange(name, property)%>% #sort by name and unit for easier viewing
  mutate(name = as.factor(name), 
         property = if_else(property == "Temperature", paste(medium, property, sep = " "), property), 
         property = if_else(str_detect(propCode, "WaterYr"), paste(property, "WaterYr", sep = " "), property),
         property = if_else(str_detect(propCode, "Discharge"), propCode, property),
         property = as.factor(property))%>%
  mutate(siteNum = as.numeric(name), 
         propNum = as.numeric(property))

nameRef <- r_df%>%
  select(name, siteNum)%>%
  distinct()%>% #Create dataframe with name references
  arrange(name)

propRef <- r_df%>%
  select(property, propNum)%>%
  distinct()%>%
  arrange(property)#Creste dataframe with property references
```

Next: view reference dataframes, write down site numbers and property numbers for export 

View name references:

```{r}
View(nameRef)
```

Enter desired site numbers below: 

```{r}
soi <- c(5,6,15,16,19) #ENTER DESIRED SITE NUMBERS HERE inside the parentheses. e.g. if you want sites 1, 3, and 19, the code should read "soi <- c(1, 3, 19)"
```

View property references:

```{r}
View(propRef)
```

Enter desired property numbers below:

```{r}
poi <- c(1,3,11,23) #ENTER DESIRED PROPERTY NUMBERS HERE inside the parentheses. e.g. if you want properties 8, 17, and 21, the code should read "soi <- c(8, 17, 21)"
```

Extract desired metadata:

```{r}
#extract desired datastreams:
toDownload <- r_df %>%
  filter(
    siteNum %in% soi&propNum %in% poi
  )

#Check to ensure all sites/data are present: 
View(toDownload)

#convert back to Python for download: 

pyDownload <- r_to_py(toDownload)
```


#### iii. Import data using a for loop

Setup:

```{python}
#Read "to download" dataframe into Python environment
pyDownload = pd.DatFrame(r.pyDownload)

#initialize empty dataframe to receive data: 

fullData = pd.DataFrame(data = None, index = None, columns = ['phenomenon_time', 'result', 'site','param', 'unit']) #initialize empty dataframe (outside loop)
```

Read in all data using a For loop - NOTE, this will take a while to run (~2-4 min per site)

```{python}
for var in range(0, len(pyDownload["uid"])): #iterate on "i" over # rows in the pyDownload DF
  print("started"+" "+pyDownload.iloc[var,1]) #print current site name for progress tracking
  data = hs_api.datastreams.get_observations(fetch_all = True, uid = pyDownload.iloc[var,0]).dataframe #read in data to temporary df
  data['site'] = pyDownload.iloc[var,1] #add site name
  data['param'] = pyDownload.iloc[var,2] #add parameter name
  data['unit'] = pyDownload.iloc[var,4] #add unit 
  fullData = pd.concat([fullData, data]) #add data to fullData df created in block above. 
  print("finished"+" "+pyDownload.iloc[var,1]) #print current site name for progress tracking
```

Convert full data file back to R for additional wrangling and analysis; replace -9999 values with NA: 

```{r}
fullData <- py_to_r(py$fullData)%>%
  rename(parameterValue = result, dateTime = phenomenon_time, parameter = param)%>% #renaming columns
  mutate(dateTime = ymd_hms(dateTime), #convert date time to POSIXct
         parameterValue = ifelse(parameterValue == -9999, NA, parameterValue)) #replace "no data" value o f-9999 with NA
```

Check date ranges for each site: 

```{r}
dateRanges <- fullData%>%
  group_by(site, parameter)%>%
  summarise(start = min(dateTime), 
            end = max(dateTime))
dateRanges
```

#### iv. Alternative formatting for DOY plotting:

Alternative data format for DOY plots: 

```{r}
dailyData <- fullData%>%
  mutate(parameter = ifelse(parameter == "Discharge_cfs", "Discharge", parameter),
    siteParam = paste(site, parameter, sep = " "))%>% #add new Site + parameter column for grouping
  mutate(doy = yday(dateTime), #make new DOY column
         date = ymd(substr(dateTime, 1,10)))%>% #make new column with just date
  group_by(siteParam, date, unit, parameter, doy)%>%
  summarise(paramDaily = mean(parameterValue, na.rm = TRUE))%>% #calculate daily average
  group_by(siteParam, doy)%>% #group by day of year
  mutate(paramMin = min(paramDaily, na.rm = T), 
         paramMax = max(paramDaily, na.rm = T), 
         paramMed = median(paramDaily, na.rm = T), 
         paramMean = mean(paramDaily, na.rm = T), 
         qt1 = quantile(paramDaily, probs = 0.1, na.rm = T)[[1]], 
         qt2 = quantile(paramDaily, probs = 0.3, na.rm = T)[[1]], 
         qt3 = quantile(paramDaily, probs = 0.7, na.rm = T)[[1]], 
         qt4 = quantile(paramDaily, probs = 0.9, na.rm = T)[[1]], )%>% #calculate day of year minimum, maximum, median, and mean, plus quantiles for each doy
  ungroup()%>% #ungroup
  arrange(siteParam, date) #sort by site+parameter and date for easier checking
```

### B. SNOTEL Data

Import data from Snotel using the snotelr package:

Get metadata for sites in the Logan River Watershed:

```{r}
snotelSites <- snotel_info(network="sntl")%>%
  filter(str_detect(description, "Logan")|str_detect(description, "Temple"))

lro_siteIds <- c(snotelSites$site_id)
```

NOTE - currently NOT including USU Doc Daniel (on drainage divide between Logan and Garden City sides) or Little Bear (S End of Cache Valley)

Download data for all sites: 

```{r}
loganSnotel <- snotel_download(site_id = lro_siteIds, network = "sntl", internal = T)%>%
  mutate(date = ymd(date), site_name= as.factor(site_name))
```


## II. Time Series Plots

### A. Discharge

Starting with a basic time series, y = discharge CFS, x = date Time, color = site. 

```{r}
#color palate: 

discharge.pal <- c("#2171B5", "#08306B",  "#9ECAE1")

#Check with updated plot: 
q.comb <- ggplot(
  filter(fullData, parameter == "Discharge_cfs"), #extract discharge data
  aes(x = dateTime, y = parameterValue, color = site) #setting mapping
)+
  geom_line()+
  theme_classic()+
  labs(x = "Date", y = "Discharge, CFS", color = "Station Name")+ #axis labels
  scale_color_manual(values = discharge.pal)+ #Setting color palate
  theme(legend.position = "bottom", 
        axis.title = element_text(size = 10, face = "bold"), 
        legend.title = element_text(size = 10, face = "bold")) #editing text appearance 

q.comb

#Save: 

ggsave("Discharge_Comb.jpg", q.comb, device = "jpeg", path = here("outputs", "plots"), units = "in", dpi = "retina", height = 6, width = 11)
```

### B. Water Temperature

Basic Time Series:

```{r}
#Initial plot:
stream.temp <- ggplot(
  filter(fullData,
         parameter == "Surface Water Temperature"), 
  aes(x = dateTime, y = parameterValue, color = site)
)+
  geom_line()+
  theme_classic()+
  labs(x = "Date", y = "Stream Temperature, deg. C", color = "Station Name")+
  scale_color_manual(values = discharge.pal)+
  theme(legend.position = "bottom", 
        axis.title = element_text(size = 10, face = "bold"), 
        legend.title = element_text(size = 10, face = "bold"))

stream.temp

#Save: 

ggsave("StreamTemp_Comb.jpg", stream.temp, device = "jpeg", path = here("outputs", "plots"), units = "in", dpi = "retina", height = 6, width = 11)
```



### C. Air Temperature

Basic time series:

```{r}
climate.pal <- c("#C7E9C0","#238B45" )

air.temp <- ggplot(
  filter(fullData,
         parameter == "Air Temperature"), 
  aes(x = dateTime, y = parameterValue, color = site)
)+
  geom_line()+
  theme_classic()+
  labs(x = "Date", y = "Air, deg. C", color = "Station Name")+
  scale_color_manual(values = climate.pal)+
  theme(legend.position = "bottom", 
        axis.title = element_text(size = 10, face = "bold"), 
        legend.title = element_text(size = 10, face = "bold"))

air.temp


#Save: 

ggsave("AirTemp_Comb.jpg", air.temp, device = "jpeg", path = here("outputs", "plots"), units = "in", dpi = "retina", height = 6, width = 11)
```

### D. Precipitation

Basic time series:

```{r}
precip <- ggplot(
  filter(fullData,
         parameter == "Precipitation"), 
  aes(x = dateTime, y = parameterValue, color = site)
)+
  geom_line()+
  theme_classic()+
  labs(x = "Date", y = "Cumulative Annual Precipitation, cm", color = "Station Name")+
  scale_color_manual(values = climate.pal)+
  theme(legend.position = "bottom", 
        axis.title = element_text(size = 10, face = "bold"), 
        legend.title = element_text(size = 10, face = "bold"))

precip


#Save: 

ggsave("Precip_Comb.jpg", precip, device = "jpeg", path = here("outputs", "plots"), units = "in", dpi = "retina", height = 6, width = 11)
```

### E. SWE

```{r}
swe <- ggplot(loganSnotel, aes(x = date, y = snow_water_equivalent, color = site_name))+
  geom_line()+
  theme_classic()+
  labs(x = "Date", y = "Snow Water Equivalent, cm", color = "Site Name")+
  scale_color_manual(values = c("darkblue", "blue", "lightblue", "cyan4"))+
  facet_wrap(~site_name)+
  theme(legend.position = "none", 
        axis.title = element_text(size = 10, face = "bold"), 
        legend.title = element_text(size = 10, face = "bold"), 
        strip.text = element_text(size = 10, face = "bold"))
swe
```


## III. Day of year plotting

Using "dailyData" dataframe created in step I. 

First, Test plotting day of year plot with one site - data prep: 

```{r}
thisyear <- format(Sys.Date(), "%Y") #Create object with this year for later filtering
test <- filter(dailyData, siteParam == "Climate Station at Franklin Basin Air Temperature") #make test dataset with one site/parameter

datemin <- year(min(test$date))
datemax <- year(max(test$date)) #Make date range object for labeling
                   
```

Plotting:

```{r}
testplot <- ggplot()+ 
  geom_line(data = test, aes(x = doy, y = paramMin, color = "Minimum"))+ #draw minimum line 
  geom_line(data = test, aes(x = doy, y = paramMed, color = paste("Median ", "(", datemin, "-", datemax, ")", sep = "")))+ #draw median line, add daterange object for label
  geom_line(data = test, aes(x = doy, y = paramMax, color = "Maximum"))+ #add maximum line
  geom_line(data = filter(test, year(date) == datemax), aes(x = doy, y = paramDaily, color = paste(datemax)))+ #filter out this year's data and add to plot
  scale_color_manual(values = c("black", "blue", "green", "red"))+ #Setting line color scale
  labs(x = "Day of Year", y = "Discharge, CFS", color = "Legend", title = unique(test$siteParam))+ #axis and legend lables
  geom_ribbon(data = test, aes(x = doy, ymax = paramMax, ymin = qt4), fill = "blue", alpha = 0.25)+
  geom_ribbon(data = test, aes(x = doy, ymax = qt4, ymin = qt3), fill = "lightblue", alpha = 0.25)+
  geom_ribbon(data = test, aes(x = doy, ymax = qt3, ymin = qt2), fill = "green", alpha = 0.25)+
  geom_ribbon(data = test, aes(x = doy, ymax = qt2, ymin = qt1), fill = "yellow", alpha = 0.25)+
  geom_ribbon(data = test, aes(x = doy, ymax = qt1, ymin = paramMin), fill = "red", alpha = 0.25)+ #adding shaded ribbons for quantiles
  theme_classic()+ #removing background formatting
  theme(legend.position = c(0.8, 0.75), #repositioning legend within plot area
        axis.text = element_text(size = 10), 
        axis.title = element_text(size = 10, face = "bold"), 
        legend.text = element_text(size = 9), 
        legend.title = element_text(size = 9, face = "bold"), 
        plot.title = element_text(size = 11, face = "bold")) #axis and legend text formatting
testplot
```

Success! Next, iterate over each site/parameter to make all plots: 

Setup: 

```{r}
allSiteParams <- unique(dailyData$siteParam) #make vector of all site/parameter combos
plot.list <- list() #empty list to recieve plots
thisyear <- format(Sys.Date(), "%Y") #Create object with this year for later filtering
q.pal <- c("#0000FF35", "#ADD8E635", "#00FF0035", "#FFFF0035" ,"#FF000035")
t.pal <- rev(q.pal)
q.line <- c("black", "blue", "green", "red")
t.line <- c("black", "red", "green", "blue") #set up color palate options
dateMins <- year(dateRanges$start)
dateMaxes <- year(dateRanges$end)
```

Make plots with a for-loop:

```{r}
for(i in 1:length(allSiteParams)){
  d <- filter(dailyData, siteParam == allSiteParams[i]) #Filter data to "ith" site/parameter

  fill.pal <- if (str_detect(allSiteParams[[i]], "Temperature")) q.pal else t.pal
  line.pal <- if (str_detect(allSiteParams[[i]], "Temperature")) t.line else q.line
  
  yLab <- paste(
    unique(d$parameter), 
    unique(d$unit), 
    sep = ", "
  )
  
  #Plotting:
  
  p <- ggplot()+ 
  geom_line(data = d, aes(x = doy, y = paramMin, color = "Minimum"))+ #draw minimum line 
  geom_line(data = d, aes(x = doy, y = paramMed, color = paste("Median ", "(", dateMins[[i]], "-", dateMaxes[[i]], ")", sep = "")))+ #draw median line, add daterange object for label
  geom_line(data = d, aes(x = doy, y = paramMax, color = "Maximum"))+ #add maximum line
  scale_color_manual(values = line.pal)+ #Setting line color scale
  labs(x = "Day of Year", y = yLab, color = "Legend", title = unique(d$siteParam), fill = "Percentiles")+ #axis and legend lables
  geom_ribbon(data = d, aes(x = doy, ymax = paramMax, ymin = qt4, fill = "90-100th"))+
  geom_ribbon(data = d, aes(x = doy, ymax = qt4, ymin = qt3, fill = "70-90th"))+
  geom_ribbon(data = d, aes(x = doy, ymax = qt3, ymin = qt2, fill = "30-70th"))+
  geom_ribbon(data = d, aes(x = doy, ymax = qt2, ymin = qt1, fill = "10-30th"))+
  geom_ribbon(data = d, aes(x = doy, ymax = qt1, ymin = paramMin, fill = "0-10th"))+ #adding shaded ribbons for quantiles
    scale_fill_manual(values = fill.pal)+
      geom_line(data = filter(d, year(date) == dateMaxes[[i]]), aes(x = doy, y = paramDaily, color = as.character(dateMaxes[[i]])), size = 0.75)+ #filter out this year's data and add to plot

    #guides(fill = "none")+
  theme_classic()+ #removing background formatting
  theme(legend.position = "bottom", #re positioning legend along bottom
        axis.text = element_text(size = 11), 
        axis.title = element_text(size = 11, face = "bold"), 
        legend.text = element_text(size = 10), 
        legend.title = element_text(size = 10, face = "bold"), 
        plot.title = element_text(size = 12, face = "bold"),
         #legend.key = element_rect(fill = alpha(fill.pal, 0.25))
        ) #axis and legend text formatting
    
  
  plot.list[[i]]<-p #Save plot as "ith" plot in plot list
  
  fn <- paste(allSiteParams[[i]], ".jpg", sep = "")
  
  ggsave(fn, p, device = "jpeg", path = here("outputs", "plots", "doy_plots"), units = "in", dpi = "retina", height = 6, width = 11)
  
  print(paste("Finished", allSiteParams[[i]], sep = " ")) #Print site name/parameter for tracking
}
```
Check outputs: 

```{r}
plot.list[[5]]
```

## IV. Basic Stats

### A. Annual Trend Analyses

Data Prep: 

```{r}
yearData <- dailyData%>%
  mutate(year = year(date))%>% #create year column
  group_by(siteParam, year, parameter, unit)%>% #group by site/parameter and year
  summarise(annualTot = sum(paramDaily, na.rm = T), 
            annualMax = max(paramDaily, na.rm = T), 
            annualMin = min(paramDaily, na.rm = T), 
            annualMean = mean(paramDaily, na.rm = T), 
            annualMed = median(paramDaily, na.rm = T), #calculate annual total, max, min, mean, and median for each site/param
            maxDate = date[which.max(paramDaily)], 
            minDate = date[which.min(paramDaily)])%>% #extract date for maximum and minimum values
  mutate(annualTot = ifelse(parameter == "Precipitation", annualMax, annualTot)) #correcting precip data (already cumulative)
```

Basic LM for annual total x year:

First, nest data based on site: 

```{r}
yearNested <- yearData %>%
  nest(data = -siteParam) #create new nested dataframe with one column for site, one column for data - data entries for each site include year and mean temp. 
```

Next, run a model for each site separately and return a summary table with the results for each site:

```{r}
yearModels <- yearNested %>%
  mutate(model = map(data, ~lm(annualTot~year, data = .)%>% #run a model of temp~year for each site (referenced by "."); store results in new column called "model"
                     tidy))%>% #tidy function to construct a summary table with model results (seperate tables for each site)
  unnest(model) %>% #unnest model column to get rows for each site, columns for model metrics
  filter(term == "year")  #remove intercept term to just get info on year for each site
yearModels
```



NOT RUN: Plotting - annual total: 


annual.plots <- list()


for(i in 1:length(allSiteParams)){
  df <- filter(yearData, siteParam == allSiteParams[i])
  
  pv <- yearModels$p.value[i]
  
  y.l <- paste("Total annual", unique(df$parameter), "," ,unique(df$unit), sep = " ")
  
  pl <- ggplot(df, aes(x = year, y = annualTot))+
    geom_point()+
    geom_smooth(se = T, method = "lm", linetype = "dashed", color = "darkgrey")+
    theme_classic()+
    labs(x = "Year", y = y.l, title = allSiteParams[[i]])+
    geom_text(aes(
      y = (0.9*max(df$annualTot, na.rm = T)), 
      x = (0.9*max(df$year, na.rm = T)),
      label = paste("P=", round(pv, 3), sep = "")))
  
  annual.plots[[i]]<-pl
  
  print(paste("Finished", allSiteParams[[i]]))
  
}

Data prep for facet plot:

```{r}
#Data prep: 

yearPlot <- merge(yearData, yearModels)

pVals <- yearPlot%>%
  group_by(siteParam)%>%
  summarize(x = mean(year), label = round(unique(p.value), 2))

climateStations <- filter(yearPlot, str_detect(siteParam, "Climate"))
streamStations <- filter(yearPlot, !str_detect(siteParam, "Climate"))

```

Plotting for climate stations:

```{r}
#Plotting - Climate Stations:

climate.plot <- ggplot(climateStations, aes(x = year, y = annualTot))+
  geom_point()+
  geom_smooth(se = T, method = "lm", linetype = "dashed", color = "darkgrey")+
  theme_classic()+
  labs(x = "Year", y = "Total annual value")+
  geom_text(data = filter(pVals, str_detect(siteParam, "Climate")), aes(x = x, label = paste("P=", label)), y = Inf, vjust = 2)+
  facet_wrap(~siteParam, scales = "free")+
  theme(axis.text = element_text(size = 11), 
        axis.title = element_text(size = 11, face = "bold"), 
        strip.text = element_text(size = 10, face = "bold"))


climate.plot

#Save plot: 

ggsave("Annual_Climate_Totals.jpg", climate.plot, device = "jpeg", path = here("outputs", "plots"), units = "in", dpi = "retina", height = 6, width = 11)
```

Plotting for stream stations: 

```{r}

stream.plot <- ggplot(streamStations, aes(x = year, y = annualTot))+
  geom_point()+
  geom_smooth(se = T, method = "lm", linetype = "dashed", color = "darkgrey")+
  theme_classic()+
  labs(x = "Year", y = "Total annual value")+
  geom_text(data = filter(pVals, !str_detect(siteParam, "Climate")), aes(x = x, label = paste("P=", label)), y = Inf, vjust = 2)+
  facet_wrap(~siteParam, scales = "free", nrow = 3, ncol = 2)+
  theme(axis.text = element_text(size = 11), 
        axis.title = element_text(size = 11, face = "bold"), 
        strip.text = element_text(size = 10, face = "bold"))


stream.plot

#Save plot:
ggsave("Annual_Stream_Totals.jpg", stream.plot, device = "jpeg", path = here("outputs", "plots"), units = "in", dpi = "retina", height = 11, width = 12)
```


### B. Individual Variable Analyses:

For each primary variable, measures of interest include: 

- **Discharge:**
  + Total annual discharge
  + Timing of peak discharge
  + Mean annual discharge
- **Stream Temperature:**
  + Annual average stream temperature
  + "Degree days" over a certain threshold - maybe 20C?
  + Average Late Summer Stream Temp (July-Aug?)
- **Precip:**
  + Total annual precip
- **Air Temperature**
  + Average annual air temperature
  + "Degree days" over a certain threshold?
  + Average seasonal temps: Winter (December-February), Spring (March-May), Summer(June-August), Fall (September-November)
- **SWE:**
  + Peak SWE
  + Date of peak SWE
  + "Meltout time": # days between peak SWE and zero SWE
  
#### i. Discharge:

First, extract data and calculate total annual discharge, mean annual discharge, and timing of peak discharge; reorganize for analysis and plotting:

```{r}
discharge <- yearData%>%
  filter(parameter == "Discharge")%>% #extract discharge data from prev. calculated annual min, mean, max, etc. 
  mutate(maxDOY =yday(maxDate))%>% #add day of year for max discharge
  select(siteParam, year, parameter, unit, annualTot, annualMean, maxDOY)%>%
  pivot_longer(!c(siteParam, year, parameter, unit), names_to = "metricName", values_to = "metricValue")
```

Nest data for modeling:  

```{r}
dischargeNest <- discharge %>%
  nest(data = -c(siteParam, metricName)) #create new nested dataframe with columns for site and metricName, one column for data - data entries for each site include year and metricValues. 
```

Run basic models for each site and metric of interest: 

```{r}
qModels <- dischargeNest %>%
  mutate(model = map(data, ~lm(metricValue~year, data = .)%>% #run a model of temp~year for each site (referenced by "."); store results in new column called "model"
                     tidy))%>% #tidy function to construct a summary table with model results (seperate tables for each site)
  unnest(model) %>% #unnest model column to get rows for each site, columns for model metrics
  filter(term == "year")  #remove intercept term to just get info on year for each site
qModels
```

Add P-values for plotting: 

```{r}
dischargePV <- merge(discharge, 
                     select(qModels,siteParam, metricName, p.value)
                     )%>%
  mutate(metricPrint = case_when(
    metricName == "annualMean"~ "Mean Annual Discharge, cfs", 
    metricName == "annualTot" ~"Total Annual Discharge, cfs", 
    metricName == "maxDOY" ~"Max. Discharge Day of Year"
  ))%>%
  mutate(site = str_remove(siteParam, "Discharge"), 
         siteMetric = paste(site, metricPrint, sep = "- "))

qPVs <- dischargePV%>%
  group_by(siteMetric, metricPrint)%>%
  summarize(x = mean(year), label = round(unique(p.value), 2))
```

Draw plots: 

```{r}
q.plot <- ggplot(dischargePV, aes(x = year, y = metricValue, color = metricPrint))+
  geom_point()+
  geom_smooth(se = T, method = "lm", linetype = "dashed")+
  theme_classic()+
  labs(x = "Year", y = "Metric Value", color = "Metric")+
  scale_color_manual(values = c("cyan4", "darkblue", "darkgreen"))+
  geom_text(data = qPVs, aes(x = x, label = paste("P=", label)), y = Inf, vjust = 2)+
  facet_wrap(~siteMetric, scales = "free", labeller = label_wrap_gen(width = 40))+
  theme(axis.text = element_text(size = 11), 
        axis.title = element_text(size = 11, face = "bold"), 
        strip.text = element_text(size = 10, face = "bold"), 
        legend.position = "bottom")


q.plot

#Save plot:
ggsave("Discharge_metrics.jpg", q.plot, device = "jpeg", path = here("outputs", "plots"), units = "in", dpi = "retina", height = 10, width = 10)
```





